### 2019-08-30 Box1.Rmd
Transfer to GitHub

---
output:
  html_document: 
    fig_caption: yes
  output: 
  pdf_document:
    fig_caption: yes
  word_document: 
    fig_caption: yes
bibliography: bibliography.bib
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggrepel)
```

**A worked example of cumulative meta-analysis to identify research waste**

We extracted the data and R code from @darras2018comparing to recreate their analysis. Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” package (@metafor) in R. The cumulative meta-analysis was ordered by publication year and plotted using the “forest” function.  To assess the point at which there is sufficient evidence and no further investigations are required we plotted the z-curve in relation to the cumulative sample size. The thresholds for significance was a z value of 1.96 or -1.96. When the z curve crosses this threshold then the level of evidence is considered sufficient. This approach (known as “trial sequential analysis”) is well developed in medicine (wetterslev2008trial). Plots were produced using ggplot2 in R (@ggplot2).

The effect size of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0 in each study (Figure 1). Trial sequential analysis shows that an evidence threshold was reached in 2015. This suggests that studies undertaken post 2015 were a waste of research resources.

```{r Make use of the code from Darras et al, message=FALSE, warning=FALSE, include=FALSE}
#Load libraries
library(readr)
library(data.table)
library(ggplot2)
library(scales)
library(cowplot)
library(metafor)
library(nlme)
library(MuMIn)
library(effects)
library(tidyverse)

#Code from the original paper (Darras et al. 2018)
#Prepare data and Run metaanalysis
#standard error function
se <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))

meta0=fread("meta-analysis - data.csv")
#give ID to all studies
meta0[,study_ID:=1:nrow(meta0)]
#remove Acevedo due to unequal sampling
meta0=meta0[first_author!="Acevedo"]
#convert to numeric
meta0=meta0[,total_time_min:=as.numeric(total_time_min)]

# Table 1 -----------------------------------------------------------------

#format column headers
meta.studies=meta0[,.(Publication=paste(first_author,year),`Sampling time (min)`=total_time_min,Microphone=microphone,`Signal to noise ratio (minimum, dB)`=signal_to_noise,`Height (cm)`=height_cm,`Number of Microphones`=number_microphones)]

# community differences ---------------------------------------------------

#compute total richness with both methods combined
meta0[,c("gamma_total_unlimited","gamma_total_identical")
      :=.(gamma_point_unlimited+unique_sound_unlimited,gamma_point_identical+unique_sound_identical)]

meta0[simultaneous_methods==1,.(common_species_unlimited=mean((gamma_point_unlimited-unique_point_unlimited)/gamma_total_unlimited,na.rm=T)
                                ,unique_point_unlimited=mean(unique_point_unlimited/gamma_total_unlimited,na.rm=T)
                                ,unique_sound_unlimited=mean(unique_sound_unlimited/gamma_total_unlimited,na.rm=T)
                                ,common_species_identical=mean((gamma_point_identical-unique_point_identical)/gamma_total_identical,na.rm=T)
                                ,unique_point_identical=mean(unique_point_identical/gamma_total_identical,na.rm=T)
                                ,unique_sound_identical=mean(unique_sound_identical/gamma_total_identical,na.rm=T))]

# calculate ROM ----------------------------------------------------------

#calculate log transformed ratio of means for alpha and gamma richness
#alpha richness identical
alpha.identical=data.table(summary(escalc(measure="ROM"
                                          ,m1i=alpha_sound_identical,m2i=alpha_point_identical
                                          ,sd1i=alpha_sound_sd_identical,sd2i=alpha_point_sd_identical
                                          ,n1i=alpha_sound_n_identical,n2i=alpha_point_n_identical
                                          ,data=meta0[!is.na(alpha_point_n_identical)])))

#alpha richness unlimited
alpha.unlimited=data.table(summary(escalc(measure="ROM"
                                          ,m1i=alpha_sound_unlimited,m2i=alpha_point_unlimited
                                          ,sd1i=alpha_sound_sd_unlimited,sd2i=alpha_point_sd_unlimited
                                          ,n1i=alpha_sound_n_unlimited,n2i=alpha_point_n_unlimited
                                          ,data=meta0[!is.na(alpha_point_n_unlimited)])))
#gamma richness identical
gamma.identical=data.table(summary(escalc(measure="ROM"
                                          ,m1i=gamma_sound_identical,m2i=gamma_point_identical
                                          ,sd1i=rep(0,nrow(meta0[!is.na(gamma_total_identical)])),sd2i=rep(0,nrow(meta0[!is.na(gamma_total_identical)]))
                                          ,n1i=total_time_min,n2i=total_time_min
                                          ,data=meta0[!is.na(gamma_total_identical)])))

#gamma richness unlimited
gamma.unlimited=data.table(summary(escalc(measure="ROM"
                                          ,m1i=gamma_sound_unlimited,m2i=gamma_point_unlimited
                                          ,sd1i=rep(0,nrow(meta0[!is.na(gamma_total_unlimited)])),sd2i=rep(0,nrow(meta0[!is.na(gamma_total_unlimited)]))
                                          ,n1i=total_time_min,n2i=total_time_min
                                          ,data=meta0[!is.na(gamma_total_unlimited)])))

#merge all ROM data together
meta1=merge(meta0,merge(alpha.identical[,.(study_ID,alpha_ROM_identical=yi,alpha_variance_ROM_identical=vi,alpha_ci.lb_identical=ci.lb,alpha_ci.ub_identical=ci.ub)]
                        ,merge(alpha.unlimited[,.(study_ID,alpha_ROM_unlimited=yi,alpha_variance_ROM_unlimited=vi,alpha_ci.lb_unlimited=ci.lb,alpha_ci.ub_unlimited=ci.ub)]
                               ,merge(gamma.unlimited[,.(study_ID,gamma_ROM_unlimited=yi,gamma_variance_ROM_unlimited=vi)]
                                      ,gamma.identical[,.(study_ID,gamma_ROM_identical=yi,gamma_variance_ROM_identical=vi)]
                                      ,all=T,by="study_ID"),all=T,by="study_ID"),all=T,by="study_ID"),all=T,by="study_ID")

# identical range (alpha) --------------------------------------------------------------

#create publication=level column for random effect
meta1[,publication:=paste(first_author,year)]

#models for identical ranges
RMA_alpha_identical=rma.mv(yi=alpha_ROM_identical,V=alpha_variance_ROM_identical
                           ,data=meta1,random=list(~1|publication))

```

```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 1. Cumulative forest plot of the meta-analysis of Darras et al. (2018) on the difference between human observers and acoustic recorders in terms of species richness.\\label{forest}"}

res <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=meta1, slab=paste(meta1$first_author, meta1$year, sep=", "))
cres<-cumul(res, order=order(meta1$year))
forest(cres)
#png("Figures/Figure1.png")
#dev.off()

```

```{r Add seq, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 2. Trial sequential analysis. The red dashed lines indicate the significance threshold (between -1.96 and 1.96). When the z-curve crosses the red line then there is sufficient evidence and no new trials are required. \\label{seq}"}

cres$total_time<-c(1360,150,40,480,7500,560,1440,2210,2210,290,570,300,600,2000,720,40,480,80,1730,100,590,4610,640,440,3140,640,600)

cumzval<-c(0.9029,
0.2495,-0.7888,
-2.0999,-1.9629,
-1.8674,-1.7743,
-1.2929,-1.0667,
-0.9152,-0.666,
-0.3349,0.1078,
1.1894,2.2064,
3.2112,4.2022,
5.5064,6.8475,
8.3989,9.6806,
11.0578,12.5002,
14.0373,15.3932,
16.9835,18.7814)

cum_sample<-c(136,              186,
              190,              238,
              363,              419,
              455,              516,
              577,              647,
              660,              720,
              780,              878,
              894,              898,
              922,              930,
              1103,              1113,
              1172,              1633,
              1666,              1710,
              1752,              1784,
              1814)

Year=c(2000,2002,2002,2002,2004,2009,2009,2009,2012,2012,2012,2012,2012,
2014,2015,2015,2015,2015,2015,2016,2016,2016,2016,2017,2017,2017,2018)

cum_df<-data.frame("z-score"=cumzval, "Cumulative n"<-cum_sample, "Year"=Year)

#png("TSA.png", width=850, height =400)
cum_df %>% 
  ggplot(aes(cum_sample,z.score))+
  labs(y="Cumulative Z Score", x="Cumulative sample size")+
  geom_point()+
  geom_line(colour="blue", linetype="dashed")+
  ylim(c(-15,15))+
  xlim(c(0,5000))+
  annotate("text", x = 1, y = 10, label = sprintf('\u2b61'),size=30) +
  annotate("text", x = 3000, y = 10, label ="Favours automatic recorder",size=4) +
  annotate("text", x = 1, y = -10, label = sprintf('\u2b63'),size=30) +
  annotate("text", x = 3000, y = -10, label ="Favours human recorder",size=4) +
  geom_hline(yintercept = 2, colour="red",linetype="dotted")+
  geom_hline(yintercept = -2, colour="red",linetype="dotted")+
  geom_text_repel(aes(label=Year),hjust=0.5, vjust=0)+
  theme_cowplot()

```

### 2019-08-30 Box1.tex
Transfer to GitHub

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\begin{document}

\begin{verbatim}
## Warning: package 'tinytex' was built under R version 3.6.1
\end{verbatim}

\begin{verbatim}
## If reinstallation fails, try install_tinytex() again. Then install the following packages:
## 
## tinytex::tlmgr_install(c("amscls", "amsfonts", "amsmath", "babel", "biber", "biblatex", "bibtex", "bigfoot", "booktabs", "caption", "cm", "dehyph", "dejavu", "dvipdfmx", "dvips", "ec", "enumitem", "environ", "etex", "etoolbox", "euenc", "fancyhdr", "fancyvrb", "filehook", "fontawesome", "fontspec", "framed", "geometry", "glyphlist", "graphics", "graphics-cfg", "graphics-def", "gsftopk", "helvetic", "hyperref", "hyphen-base", "ifluatex", "ifmtarg", "iftex", "ifxetex", "inconsolata", "knuth-lib", "kpathsea", "l3kernel", "l3packages", "latex", "latex-bin", "latex-fonts", "latexconfig", "latexmk", "lm", "lm-math", "logreq", "lualibs", "luaotfload", "luatex", "makeindex", "mathspec", "metafont", "mfware", "microtype", "ms", "natbib", "oberdiek", "parskip", "pdftex", "pgf", "plain", "scheme-infraonly", "setspace", "sourcesanspro", "tcolorbox", "tetex", "tex", "tex-ini-files", "texlive.infra", "times", "tipa", "titling", "tlgs", "tlperl", "tlpsv", "tools", "trimspaces", "unicode-data", "unicode-math", "upquote", "url", "xargs", "xcolor", "xetex", "xetexconfig", "xifthen", "xkeyval", "xpatch", "xunicode", "zapfding"))
\end{verbatim}

\begin{verbatim}
## The directory C:\Users\matthew.grainger\AppData\Roaming\TinyTeX/texmf-local is not empty. It will be backed up to C:\Users\MATTHE~1.GRA\AppData\Local\Temp\RtmpG2F7Ge\file2c1832d17a7a and restored later.
\end{verbatim}

\begin{verbatim}
## tlmgr conf auxtrees remove "C:/Programs/R/R-36~1.0/share/texmf"
\end{verbatim}

\begin{verbatim}
## tlmgr path remove
\end{verbatim}

\begin{verbatim}
## Starting to install TinyTeX to C:\Users\matthew.grainger\AppData\Roaming\TinyTeX. It will take a few minutes.
\end{verbatim}

\begin{verbatim}
## Next you may see two error dialog boxes about the missing luatex.dll, and an error message like "Use of uninitialized value in bitwise or (|)..." in the end. These messages can be ignored.
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names = TRUE), target,
## recursive = TRUE): problem copying TinyTeX\tlpkg\installer\tar.exe to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\installer\tar.exe:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\bin\libgcc_s_dw2-1.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\bin\libgcc_s_dw2-1.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\bin\libstdc++-6.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\bin\libstdc+
## +-6.dll: Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\bin\libwinpthread-1.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\bin\libwinpthread-1.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names = TRUE), target,
## recursive = TRUE): problem copying TinyTeX\tlpkg\tlperl\bin\perl.exe to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\bin\perl.exe:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem
## copying TinyTeX\tlpkg\tlperl\bin\perl528.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\bin\perl528.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Compress\Raw\Zlib\Zlib.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Compress\Raw\Zlib\Zlib.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Cwd\Cwd.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Cwd\Cwd.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Digest\MD5\MD5.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Digest\MD5\MD5.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Digest\SHA\SHA.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Digest\SHA\SHA.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Encode\Encode.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Encode\Encode.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Fcntl\Fcntl.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Fcntl\Fcntl.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\File\Glob\Glob.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\File\Glob\Glob.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\IO\IO.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\IO\IO.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\List\Util\Util.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\List\Util\Util.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\POSIX\POSIX.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\POSIX\POSIX.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Socket\Socket.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Socket\Socket.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Storable\Storable.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Storable\Storable.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Time\HiRes\HiRes.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Time\HiRes\HiRes.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Win32\Win32.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Win32\Win32.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\lib\auto\Win32API\File\File.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\lib\auto\Win32API\File\File.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\site\lib\auto\Win32\API\API.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\site\lib\auto\Win32\API\API.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\site\lib\auto\Win32\OLE\OLE.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\site\lib\auto\Win32\OLE\OLE.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\site\lib\auto\Win32\Shortcut\Shortcut.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\site\lib\auto\Win32\Shortcut\Shortcut.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## Warning in file.copy(list.files("TinyTeX", full.names
## = TRUE), target, recursive = TRUE): problem copying
## TinyTeX\tlpkg\tlperl\site\lib\auto\Win32API\Registry\Registry.dll to C:
## \Users\matthew.grainger\AppData\Roaming\TinyTeX\tlpkg\tlperl\site\lib\auto\Win32API\Registry\Registry.dll:
## Permission denied
\end{verbatim}

\begin{verbatim}
## TinyTeX installed to C:\Users\matthew.grainger\AppData\Roaming\TinyTeX
\end{verbatim}

\begin{verbatim}
## Please quit and reopen your R session and IDE (if you are using one, such as RStudio or Emacs) and check if tinytex:::is_tinytex() is TRUE.
\end{verbatim}

\textbf{A worked example of cumulative meta-analysis to identify
research waste}

As an example of the approach we can look at the potential of autonomous
acoustic recorders to replace human observers in wildlife sampling and
monitoring, which now has a long history in the ecological literature
(e.g.~Ralph et al.~1998). Technological advances over the last two
decades have allowed this potential to be explored fully. Well over 150
field studies have been carried out that address this issue either
directly or indirectly and seek to address the question of whether
acoustic recorders can replace human observers in wildlife surveys. In
2018, Darras et al.~explored the pooled effect of these types of studies
using a meta-analysis. They concluded that when human observers (using
point counts) and sound recorders sample areas of equal size then there
is no difference between estimates of bird species richness.

By using a cumulative meta-analysis we can answer the question ``Do we
need another study quantifying the difference between acoustic recorders
and human observers for bird survey point counts?'' In addition, we can
show the historical point at which studies into this topic could have
been stopped and research waste (in a restricted sense) could have been
avoided.

We extracted the data and R code from Darras et al.~(2018) to recreate
their analysis. Building on their random effects meta-analysis we ran a
cumulative meta-analysis using the ``cumul'' function in the ``metafor''
package (Viechtbauer 2010) in R. The cumulative meta-analysis was
ordered by publication year and plotted using the ``forest'' function.
To assess the point at which there is sufficient evidence and no further
investigations are required we plotted the z-curve in relation to the
cumulative sample size. The thresholds for significance was a z value of
1.96 or -1.96. When the z curve crosses this threshold then the level of
evidence is considered sufficient. This approach (known as ``trial
sequential analysis'') is well developed in medicine (Wetterslev et
al.~2008). Plots were produced using ggplot2 in R (Wickham 2016).

The effect size of studies investigating the difference between
autonomous acoustic recorders and human observers in terms of bird
species richness estimates was consistently close to 0 in each study
(Figure 1). Trial sequential analysis shows that an evidence threshold
was reached in 2015. This suggests that studies undertaken post 2015
were a waste of research resources.

\begin{figure}
\centering
\includegraphics{Figures/Figure1.png}
\caption{Figure caption \label{figurelabel}}
\end{figure}

\end{document}

### 2019-08-30 MS.Rmd
Transfer to GitHub

--- 2019-08-30 Box1.Rmd
@@
 ---
+title: "Research waste in applied ecology and conservation"
+csl: nature.csl
 output:
-  html_document: 
-    fig_caption: yes
-  output: 
-  pdf_document:
-    fig_caption: yes
-  word_document: 
-    fig_caption: yes
+  html_document: default
+  word_document: default
+  pdf_document: default
 bibliography: bibliography.bib
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-library(ggrepel)
+```{r setup, include=FALSE}
+
+knitr::opts_chunk$set(echo = TRUE)
+
 ```
+
+Matthew J. Grainger^1^*, Frederike C. Bolam^2^, Gavin B. Stewart^2^, Erlend B. Nilsen^1^
+
+^1^Norwegian Institute for Nature Research, P.O. Box 5685 Torgarden, 7485, Trondheim, Norway
+
+^2^Modelling Evidence and Policy Research Group, School of Natural and Environmental Sciences, Ridley Building 2, Claremont Road, Newcastle upon Tyne NE1 7RU, UK
+
+*corresponding author: matthew.grainger\@nina.no
+
+“Research waste” is a well established concept in medical research (@glasziou2018research).
+Research is wasted when its outcomes cannot be used for the benefit of society (@chalmers2009avoidable).
+Waste can occur at any of the four stages of research (as defined by @chalmers2009avoidable); question setting; methods; accessibility; and reporting.
+In recent years there has been an increased focus in ecology on methods (e.g.
+@fraser2018questionable; @nilsen_bowler_linnell_2019) which have called for methodological improvements in individual studies.
+Open science is fast becoming mainstream in ecological publishing which is vastly improving accessibility (@powers2019open).
+Evidence synthesis approaches such as systematic review and meta-analysis have become stalwarts of ecology and conservation research (@gurevitch2018meta) leading to improved reporting for the benefit of decision-makers.
+Despite these advances research waste still occurs in ecology and here we focus on a critical area where ecology and conservation can make large steps forward in research waste reduction.
+
+Chalmers & Glasziou (2009) separate what we suggest is part of the same research stage across the reporting and question setting stages.
+Namely, that new studies do not take in to account the body of evidence that have taken place prior to them (categorised as reporting by @chalmers2009avoidable) and that low priority questions are being asked.
+A low priority question may be one that is not relevant to stakeholders (and @chalmers2009avoidable suggest that stakeholders be apart of the research commissioning process).
+Alternatively a low priority question may be one that has already been sufficeintly addressed in the literature so that any further studies are a waste of resources and should be avoided.
+
+In medicine, one way in which this particular type of research waste has been identified is by looking at cumulative meta-analysis.
+A cumulative meta-analysis typically describes the accumulation of evidence (e.g., about the effectiveness of an intervention) as the available estimates are added to the analysis in chronological order (@lau1992cumulative).
+By carrying out this approach a researcher or funder can show the point at which there is sufficient evidence to be confident that a reported effect is true and thus new trials are no longer required.
+
+As an example of the approach we can look at the potential of autonomous acoustic recorders to replace human observers in wildlife sampling and monitoring, which now has a long history in the ecological literature (starting from at least @ralph1995monitoring).
+Technological advances over the last two decades have allowed this potential to be explored fully.
+Well over 150 field studies have been carried out that address this issue either directly or indirectly and seek to address the question of whether acoustic recorders can replace human observers in wildlife surveys.
+Darras et al.
+(2018) explored the pooled effect of these types of studies using a meta-analysis.
+They concluded that when human observers (using point counts) and sound recorders sample areas of equal size then there is no difference between estimates of bird species richness.
+Despite this, using their search terms in a Web of Science search identifies another 14 papers that have been published after Darras et al.
+(2018) was submitted.
+By using a cumulative meta-analysis we can answer the question “Do we need another study quantifying the difference between acoustic recorders and human observers for bird survey point counts?” In addition, we can show the historical point at which studies into this topic could have been stopped and research waste (in a restricted sense) could have been avoided.
 
@@
 When the z curve crosses this threshold then the level of evidence is considered sufficient.
-This approach (known as “trial sequential analysis”) is well developed in medicine (wetterslev2008trial).
+This approach (known as “trial sequential analysis”) is well developed in medicine (@wetterslev2008trial).
 Plots were produced using ggplot2 in R (@ggplot2).
@@
 library(tidyverse)
+library(ggrepel)
 
@@
 
-```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 1.
-Cumulative forest plot of the meta-analysis of Darras et al.
+```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
 (2018) on the difference between human observers and acoustic recorders in terms of species richness.\\label{forest}"}
@@
 cres<-cumul(res, order=order(meta1$year))
+png("Figures/Figure1.png")
 forest(cres)
-#png("Figures/Figure1.png")
-#dev.off()
+
+dev.off()
 
@@
 
-```{r Add seq, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 2.
-Trial sequential analysis.
+```{r Add seq, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Trial sequential analysis.
 The red dashed lines indicate the significance threshold (between -1.96 and 1.96).
@@
 ```
+
+To reduce research waste we need to be able to first quantify it.
+One option is to use cumulative meta-analysis.
+The approach demonstrated here is well established in the medical literature.
+It is well known and tested.
+Therefore, it should not be challenging to integrate in to conservation and applied ecology workflows.
+Ideally, funding decisions would take into account the available evidence before giving grants to well studied topics.
+
+There are at least two important caveats that need to be discussed.
+The first is that currently, due to heterogeneity in reporting as well as the drive for novelty in publications that meta-analysis is challenging in applied ecology as it is.
+There might not be sufficient good quality research to quantify the cumulative effect of even some apparently well studied phenomena.
+In cases such as this one might consider using a systems modelling approach to identify key areas of uncertainty in a topic that could be prioritised for research focus.
+
+It is important to remember that research funding for fieldwork might not only be about answering a research question and could be more targeted to the development of researchers skills (PhD students, etc.) in their field of interest.
+Funding for training and skill development should be given weight in any funding prioritisation.
+
+“We need less research, better research, and research done for the right reasons” (Altmann, 1994).
+
+# References

### 2019-08-30 Rcode.R

@@ -1,0 +1,192 @@
+#Load libraries
+library(readr)
+library(data.table)
+library(ggplot2)
+library(scales)
+library(cowplot)
+library(metafor)
+library(nlme)
+library(MuMIn)
+library(effects)
+library(tidyverse)
+library(ggsci)
+library(gganimate)
+#Code from the original paper (Darras et al. 2018)
+#Prepare data and Run metaanalysis
+#standard error function
+se <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
+
+meta0=fread("meta-analysis - data.csv")
+#give ID to all studies
+meta0[,study_ID:=1:nrow(meta0)]
+#remove Acevedo due to unequal sampling
+meta0=meta0[first_author!="Acevedo"]
+#convert to numeric
+meta0=meta0[,total_time_min:=as.numeric(total_time_min)]
+
+# Table 1 -----------------------------------------------------------------
+
+#format column headers
+meta.studies=meta0[,.(Publication=paste(first_author,year),`Sampling time (min)`=total_time_min,Microphone=microphone,`Signal to noise ratio (minimum, dB)`=signal_to_noise,`Height (cm)`=height_cm,`Number of Microphones`=number_microphones)]
+
+# community differences ---------------------------------------------------
+
+#compute total richness with both methods combined
+meta0[,c("gamma_total_unlimited","gamma_total_identical")
+      :=.(gamma_point_unlimited+unique_sound_unlimited,gamma_point_identical+unique_sound_identical)]
+
+meta0[simultaneous_methods==1,.(common_species_unlimited=mean((gamma_point_unlimited-unique_point_unlimited)/gamma_total_unlimited,na.rm=T)
+                                ,unique_point_unlimited=mean(unique_point_unlimited/gamma_total_unlimited,na.rm=T)
+                                ,unique_sound_unlimited=mean(unique_sound_unlimited/gamma_total_unlimited,na.rm=T)
+                                ,common_species_identical=mean((gamma_point_identical-unique_point_identical)/gamma_total_identical,na.rm=T)
+                                ,unique_point_identical=mean(unique_point_identical/gamma_total_identical,na.rm=T)
+                                ,unique_sound_identical=mean(unique_sound_identical/gamma_total_identical,na.rm=T))]
+
[…]

### 2019-08-30 Manuscript/MS.Rmd
Clean out old files

--- 2019-08-30 Box1.Rmd
@@
 ---
-output:
-  html_document: 
-    fig_caption: yes
+title: "Research waste in applied ecology and conservation"
+csl: nature.csl
 output:
@@
     fig_caption: yes
+  html_document: 
+    fig_caption: yes
 bibliography: bibliography.bib
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-library(ggrepel)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+
 ```
 
-**A worked example of cumulative meta-analysis to identify research waste**
+Matthew J. Grainger^1^*, Frederike C. Bolam^2^, Gavin B. Stewart^2^, Erlend B. Nilsen^1^
+
+^1^Norwegian Institute for Nature Research, P.O. Box 5685 Torgarden, 7485, Trondheim, Norway
+
+^2^Modelling Evidence and Policy Research Group, School of Natural and Environmental Sciences, Ridley Building 2, Claremont Road, Newcastle upon Tyne NE1 7RU, UK
+
+*corresponding author: matthew.grainger\@nina.no
+
+*Wider adoption of methods such as cumulative meta-analysis is vital to reduce research waste in ecology and conservation.
+A formal assessment of the accumulated knowledge prior to research approval would reduce the waste in already limited resources caused by asking questions of low priority.
+Researchers, funders and decision-makers all share the responsibility to reduce research waste.*
+
+“Research waste” is a well established concept in medical research @glasziou2018research.
+Research is wasted when its outcomes cannot be used for the benefit of society @chalmers2009avoidable.
+Waste can occur at any of the four stages of research @chalmers2009avoidable; question setting; methods; accessibility; and reporting.
+In recent years there has been an increased focus in ecology on methods (e.g.
+@fraser2018questionable; @nilsen_bowler_linnell_2019) which have called for methodological improvements in individual studies.
+Open science is fast becoming mainstream in ecological publishing which is vastly improving accessibility @powers2019open.
+Evidence synthesis approaches such as systematic review and meta-analysis have become stalwarts of ecology and conservation research @gurevitch2018meta leading to improved reporting for the benefit of decision-makers.
+Despite these advances research waste still occurs in ecology and here we focus on the critical area of question setting where ecology and conservation can make large steps forward in research waste reduction.
+
+The question setting stage has two related areas where research waste can be reduced.
+Namely,  new studies do not take into account the body of evidence that have taken place prior to them and as a consequence low priority questions are being asked.
+A low priority question may be one that is not relevant to stakeholders (to remedy this stakeholders should be apart of the research commissioning process @chalmers2009avoidable), or be one that has already been sufficiently addressed in the literature so that we already know the outcome with high certainty.
+Any further studies could then be considered a waste of resources, because they will not add (or only marginally so) to our knowledge base.
+
+In medicine, one way in which this particular type of research waste has been identified is by conducting cumulative meta-analysis based on effect sizes from previous studies.
+A cumulative meta-analysis typically describes the accumulation of evidence (e.g., about the effectiveness of an intervention) across time, and  available estimates are added to the analysis in chronological order @lau1992cumulative.
+Using cumulative meta-analysis, a researcher,  funding agency or decision maker can identify the point at which there is sufficient evidence to be confident that a reported effect is true.
+At this stage new trials are no longer required to predict the outcome with satisfactory certainty.
+
+As an example of the approach applied to an applied ecology situation we can look at the potential of autonomous acoustic recorders to replace human observers in wildlife sampling and monitoring, which now has a long history in the ecological literature (starting from at least @ralph1995monitoring).
+Technological advances over the last two decades have allowed this potential to be explored fully.
+Well over 150 field studies have been carried out that address this issue either directly or indirectly and seek to address the question of whether acoustic recorders can replace human observers in wildlife surveys.
+A meta-analysis in 2018 @darras2018comparing
+ explored the pooled effect of these studies using a meta-analytic approach to estimate species richness of birds.
+Based on the combined evidence from the included studies, they concluded that when human observers (using point counts) and sound recorders sample areas of equal size then there is no difference between estimates of bird species richness.
+When properly conducted (see specific advice in @darras2018comparing), it can be inferred that sound recorders can be used to monitor aspects of biodiversity as efficiently as human observers.
+Twenty-eight primary studies published between 2000 and 2017 were included in their meta-analysis.
+
+Taking the role as a research funder or researcher at the question setting stage, it would be nice if we could know if we need another study quantifying the difference between acoustic recorders and human observers for bird survey point counts to make this conclusion?
+This is exactly the situation where the cumulative meta-analysis approach is instrumental.
 
 We extracted the data and R code from @darras2018comparing to recreate their analysis.
-Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” package (@metafor) in R.
+Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” @metafor
+package in R.
 The cumulative meta-analysis was ordered by publication year and plotted using the “forest” function.
@@
 When the z curve crosses this threshold then the level of evidence is considered sufficient.
-This approach (known as “trial sequential analysis”) is well developed in medicine (wetterslev2008trial).
-Plots were produced using ggplot2 in R (@ggplot2).
+This approach (known as “trial sequential analysis”) is well developed in medicine @wetterslev2008trial.
+Plots were produced using ggplot2 @ggplot2.
 
-The effect size of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0 in each study (Figure 1).
-Trial sequential analysis shows that an evidence threshold was reached in 2015.
-This suggests that studies undertaken post 2015 were a waste of research resources.
+The effect size (i.e the magnitude of the difference between intervention and control) of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0 in each study (Figure 1).
+This means that there was no clear difference between acoustic recorders and human observers on bird point counts.
+Trial sequential analysis (Figure 2) shows that the significance threshold (between -1.96 and 1.96) was crossed in 2015 favouring automatic recorders over human observers in bird point surveys.
+This suggests that studies undertaken post 2015 were a “waste” of research resources.
+
+To reduce research waste we need to be able to first quantify it.
+One option is to use cumulative meta-analysis.
+The approach demonstrated here is well established in the medical literature.
+It is well known and tested.
+Therefore, it should not be challenging to integrate in to conservation and applied ecology workflows.
+Ideally, funding decisions will take into account the available evidence before giving grants to well studied topics.
+
+More frequent application of systematic reviews combined with meta-analysis in general, and cumulative meta-analysis in particular, would be one important remedy to reduce research waste in ecology and conservation.
+However, there are at least two important caveats that need to be highlighted.
+First, is that currently, due to heterogeneity in reporting as well as the drive for novelty in publications that meta-analysis is challenging in applied ecology.
+There might not be sufficient good quality research to quantify the cumulative effect of even some apparently well studied phenomena.
+In cases such as this one might consider using a systems modelling approach to identify key areas of uncertainty in a topic that could be prioritised for research focus.
+
+It is also important to remember that research funding might not only be about answering a research question and could be more targeted to the development of researchers skills (PhD students, etc.) or to engaging with stakeholders.
+Funding for training, skill development and outreach activities should be given appropriate weight in any funding prioritisation.
+
+Research waste can be reduced and it is the responsibility of funders as well as individual researchers to do so.
+We agree with the statement targeted at medicine 25 years ago that “We need less research, better research, and research done for the right reasons” @altman1994scandal.
+Without a change in focus ecology and conservation funding will continue to be wasted which will be detrimental to our efforts to provide solutions to global challenges.
+
+# References {-}
+
+<div id="refs"></div>
+
+\newpage
 
@@
 library(tidyverse)
-
+library(ggsci)
+library(gganimate)
 #Code from the original paper (Darras et al. 2018)
@@
 
-```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 1.
-Cumulative forest plot of the meta-analysis of Darras et al.
+```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
 (2018) on the difference between human observers and acoustic recorders in terms of species richness.\\label{forest}"}
-
+### Add cumulative meta-analysis
 res <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=meta1, slab=paste(meta1$first_author, meta1$year, sep=", "))
 cres<-cumul(res, order=order(meta1$year))
-forest(cres)
-#png("Figures/Figure1.png")
-#dev.off()
-
+forest(cres, col="blue")
 ```
 
-```{r Add seq, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 2.
-Trial sequential analysis.
+```{r Add seq, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Trial sequential analysis.
 The red dashed lines indicate the significance threshold (between -1.96 and 1.96).
@@
 
-cres$total_time<-c(1360,150,40,480,7500,560,1440,2210,2210,290,570,300,600,2000,720,40,480,80,1730,100,590,4610,640,440,3140,640,600)
+cumzval<-cumsum(cres$zval)
+cum_sample<-cumsum(meta1$alpha_point_n_identical)  
+# cum_sample<-c(136,              186,
+#               190,              238,
+#               363,              419,
+#               455,              516,
+#               577,              647,
+#               660,              720,
+#               780,              878,
+#               894,              898,
+#               922,              930,
+#               1103,              1113,
+#               1172,              1633,
+#               1666,              1710,
+#               1752,              1784,
+#               1814)
 
-cumzval<-c(0.9029,
-0.2495,-0.7888,
--2.0999,-1.9629,
--1.8674,-1.7743,
--1.2929,-1.0667,
--0.9152,-0.666,
--0.3349,0.1078,
-1.1894,2.2064,
-3.2112,4.2022,
-5.5064,6.8475,
-8.3989,9.6806,
-11.0578,12.5002,
-14.0373,15.3932,
-16.9835,18.7814)
+Year=meta1$year
 
-cum_sample<-c(136,              186,
-              190,              238,
-              363,              419,
-              455,              516,
-              577,              647,
-              660,              720,
-              780,              878,
-              894,              898,
-              922,              930,
-              1103,              1113,
-              1172,              1633,
-              1666,              1710,
-              1752,              1784,
-              1814)
+Year=gsub("a","",Year)
+Year=gsub("b","",Year)
+Year<-as.integer(Year)
+Year<-Year[order(Year)]
+# Year=c(2000,2002,2002,2002,2004,2009,2009,2009,2012,2012,2012,2012,2012,
+# 2014,2015,2015,2015,2015,2015,2016,2016,2016,2016,2017,2017,2017,2018)
+library(ggrepel)
+cum_df<-data.frame("z-score"=cumzval, "Cumulative n"=cum_sample, "Year"=Year)
 
-Year=c(2000,2002,2002,2002,2004,2009,2009,2009,2012,2012,2012,2012,2012,
-2014,2015,2015,2015,2015,2015,2016,2016,2016,2016,2017,2017,2017,2018)
-
-cum_df<-data.frame("z-score"=cumzval, "Cumulative n"<-cum_sample, "Year"=Year)
-
+cum_df2<-cum_df[c(1,18,27),]
 #png("TSA.png", width=850, height =400)
 cum_df %>% 
-  ggplot(aes(cum_sample,z.score))+
-  labs(y="Cumulative Z Score", x="Cumulative sample size")+
-  geom_point()+
-  geom_line(colour="blue", linetype="dashed")+
-  ylim(c(-15,15))+
-  xlim(c(0,5000))+
-  annotate("text", x = 1, y = 10, label = sprintf('\u2b61'),size=30) +
-  annotate("text", x = 3000, y = 10, label ="Favours automatic recorder",size=4) +
-  annotate("text", x = 1, y = -10, label = sprintf('\u2b63'),size=30) +
-  annotate("text", x = 3000, y = -10, label ="Favours human recorder",size=4) +
-  geom_hline(yintercept = 2, colour="red",linetype="dotted")+
-  geom_hline(yintercept = -2, colour="red",linetype="dotted")+
-  geom_text_repel(aes(label=Year),hjust=0.5, vjust=0)+
-  theme_cowplot()
+  ggplot(aes(cum_sample,z.score, colour=as.factor(Year)))+
+  scale_fill_npg()+
+  labs(y="Cumulative z score", x="Cumulative sample size")+
+  geom_point(size=6)+
+  geom_line(colour="blue", linetype="dashed", size=1)+
+  ylim(c(-20,20))+
+  xlim(c(0,2000))+
 
+  #annotate("text", x = 1, y = 10, label = sprintf('\u2b61'),size=30) +
+  #annotate("text", x = 3000, y = 10, label ="Favours automatic recorder",size=4) +
+  #annotate("text", x = 1, y = -10, label = sprintf('\u2b63'),size=30) +
+  #annotate("text", x = 3000, y = -10, label ="Favours human recorder",size=4) +
+  geom_text(data=cum_df2, aes(Cumulative.n,z.score,label=as.factor(Year)),hjust=1.3, vjust=1, colour="black")+
+
+  #geom_text(aes(label=as.factor(Year)),hjust=0.5, vjust=0)+
+  geom_hline(yintercept = 2, colour="red",linetype="dotted", size=1.6)+
+  geom_hline(yintercept = -2, colour="red",linetype="dotted", size=1.6)+
+  theme_classic() +
+  theme(axis.line = element_line(size = 1.6, color = rgb(0,0,0,max=255)))+
+  theme(legend.position = "none")+
+  theme(text=element_text(size=16,  family="serif"))
 ```

### 2019-08-30 Rcode.R

@@ -11,11 +11,15 @@ library(effects)
 library(tidyverse)
 library(ggsci)
 library(gganimate)
-#Code from the original paper (Darras et al. 2018)
+library(ggrepel)
+#################################################################################################################
+################################Code from the original paper (Darras et al. 2018)################################
+#################################################################################################################
 #Prepare data and Run metaanalysis
+
 #standard error function
 se <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
-
+#Get data
 meta0=fread("meta-analysis - data.csv")
 #give ID to all studies
 meta0[,study_ID:=1:nrow(meta0)]
@@ -23,14 +27,9 @@ meta0[,study_ID:=1:nrow(meta0)]
 meta0=meta0[first_author!="Acevedo"]
 #convert to numeric
 meta0=meta0[,total_time_min:=as.numeric(total_time_min)]
-
-# Table 1 -----------------------------------------------------------------
-
 #format column headers
 meta.studies=meta0[,.(Publication=paste(first_author,year),`Sampling time (min)`=total_time_min,Microphone=microphone,`Signal to noise ratio (minimum, dB)`=signal_to_noise,`Height (cm)`=height_cm,`Number of Microphones`=number_microphones)]
-
 # community differences ---------------------------------------------------
-
 #compute total richness with both methods combined
 meta0[,c("gamma_total_unlimited","gamma_total_identical")
       :=.(gamma_point_unlimited+unique_sound_unlimited,gamma_point_identical+unique_sound_identical)]
@@ -89,52 +88,31 @@ meta1[,publication:=paste(first_author,year)]
 RMA_alpha_identical=rma.mv(yi=alpha_ROM_identical,V=alpha_variance_ROM_identical
                            ,data=meta1,random=list(~1|publication))
 
-
+#################################################################################################################
+#################################################################################################################
 
 ### Add cumulative meta-analysis
[…]

### 2019-08-30 Rcode.R
update Markdown

@@ -1,17 +1,15 @@
 #Load libraries
-library(readr)
-library(data.table)
-library(ggplot2)
-library(scales)
-library(cowplot)
-library(metafor)
-library(nlme)
-library(MuMIn)
-library(effects)
-library(tidyverse)
-library(ggsci)
-library(gganimate)
-library(ggrepel)
+# List of packages for session
+.packages = c("readr", "data.table", "scales", "cowplot", "metafor",
+"nlme", "MuMIn", "effects", "tidyverse", "ggsci", "gganimate", "ggrepel")
+
+# Install CRAN packages (if not already installed)
+.inst <- .packages %in% installed.packages()
+if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
+
+# Load packages into session 
+lapply(.packages, require, character.only=TRUE)
+
 #################################################################################################################
 ################################Code from the original paper (Darras et al. 2018)################################
 #################################################################################################################
@@ -20,7 +18,7 @@ library(ggrepel)
 #standard error function
 se <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
 #Get data
-meta0=fread("meta-analysis - data.csv")
+meta0=fread("data/meta-analysis - data.csv")
 #give ID to all studies
 meta0[,study_ID:=1:nrow(meta0)]
 #remove Acevedo due to unequal sampling

### 2019-09-18 Flowchart.R
Add flowchart

@@ -1,0 +1,70 @@
+library(DiagrammeR)
+grViz("digraph flowchart {
+      rankdir=LR
+      # node definitions with substituted label text
+      node [fontname = Helvetica, shape = rectangle]        
+      tab1 [label = '@@1']
+      tab2 [label = '@@2']
+      tab3 [label = '@@3']
+      tab4 [label = '@@4']
+      tab5 [label = '@@5']
+
+      # edge definitions with the node IDs
+      tab1 -> tab2 -> tab3 -> tab4;
+      tab1->tab5;
+      tab2->tab5;
+      tab3->tab5;
+      tab4->tab5;
+      
+      }
+
+      [1]: 'Questions relevant to stakeholders'
+      [2]: 'Appropriate design & methods'
+      [3]: 'Accessible full publication'
+      [4]: 'Unbiased reporting'
+      [5]: 'Research waste'
+      ")
+
+
+grViz("
+digraph boxes_and_circles {
+
+  # a 'graph' statement
+  graph [overlap = true, fontsize = 10]
+
+  # several 'node' statements
+  node [shape = box,
+        fontname = Helvetica]
+      tab1 [label = '@@1']
+      tab2 [label = '@@2']
+      tab3 [label = '@@3']
+      tab4 [label = '@@4']
+      tab5 [label = '@@5']
+  node [shape = circle,
+        fixedsize = true,
+        width = 2] // sets as circles
+    tab6 [label = '@@6']
+    node [shape = diamond, width=3.7]
+    tab7 [label = '@@7']
+  # several 'edge' statements
+ tab1->tab2->tab3->tab4
+ tab1->tab5
+  tab2->tab5
+   tab3->tab5
+    tab4->tab5
+    tab4->tab6 
+    tab6->tab7 [style=dotted]
+          tab6->tab5 [style=dashed]
+    tab7->tab1 [style=dotted]
+        tab7->tab2 [style=dotted]
+    
+    
+}
+      [1]: 'Questions relevant to stakeholders'
+      [2]: 'Appropriate design & methods'
+      [3]: 'Accessible full publication'
+      [4]: 'Unbiased reporting'
+      [5]: 'Research waste'
+      [6]: 'Evidence synthesis'
+      [7]: 'Cummulative meta-analysis'
+")

### 2019-09-19 Flowchart.R
Add colour outline to flowbox

@@ -1,29 +1,29 @@
 library(DiagrammeR)
-grViz("digraph flowchart {
-      rankdir=LR
-      # node definitions with substituted label text
-      node [fontname = Helvetica, shape = rectangle]        
-      tab1 [label = '@@1']
-      tab2 [label = '@@2']
-      tab3 [label = '@@3']
-      tab4 [label = '@@4']
-      tab5 [label = '@@5']
-
-      # edge definitions with the node IDs
-      tab1 -> tab2 -> tab3 -> tab4;
-      tab1->tab5;
-      tab2->tab5;
-      tab3->tab5;
-      tab4->tab5;
-      
-      }
-
-      [1]: 'Questions relevant to stakeholders'
-      [2]: 'Appropriate design & methods'
-      [3]: 'Accessible full publication'
-      [4]: 'Unbiased reporting'
-      [5]: 'Research waste'
-      ")
+# grViz("digraph flowchart {
+#       rankdir=LR
+#       # node definitions with substituted label text
+#       node [fontname = Helvetica, shape = rectangle]        
+#       tab1 [label = '@@1']
+#       tab2 [label = '@@2']
+#       tab3 [label = '@@3']
+#       tab4 [label = '@@4']
+#       tab5 [label = '@@5']
+# 
+#       # edge definitions with the node IDs
+#       tab1 -> tab2 -> tab3 -> tab4;
+#       tab1->tab5;
+#       tab2->tab5;
+#       tab3->tab5;
+#       tab4->tab5;
+#       
+#       }
+# 
+#       [1]: 'Questions relevant to stakeholders'
+#       [2]: 'Appropriate design & methods'
+#       [3]: 'Accessible full publication'
+#       [4]: 'Unbiased reporting'
+#       [5]: 'Research waste'
+#       ")
 
 
 grViz("
@@ -35,29 +35,25 @@ digraph boxes_and_circles {
   # several 'node' statements
   node [shape = box,
         fontname = Helvetica]
-      tab1 [label = '@@1']
-      tab2 [label = '@@2']
-      tab3 [label = '@@3']
-      tab4 [label = '@@4']
-      tab5 [label = '@@5']
+      tab1 [label = '@@1',color = blue]
+      tab2 [label = '@@2',color = blue]
+      tab3 [label = '@@3',color = blue]
+      tab4 [label = '@@4',color = blue]
+      tab5 [label = '@@5',color = red]
[…]

### 2019-09-20 Manuscript/MS.Rmd
update markdown ms

@@
 ---
-title: "Research waste in applied ecology and conservation"
+title: "Maximising the leverage of existing knowledge could reduce research waste in applied ecology and conservation"
 csl: nature.csl
@@
 
-*Wider adoption of methods such as cumulative meta-analysis is vital to reduce research waste in ecology and conservation.
-A formal assessment of the accumulated knowledge prior to research approval would reduce the waste in already limited resources caused by asking questions of low priority.
-Researchers, funders and decision-makers all share the responsibility to reduce research waste.*
+*Where research does not lead to benefits to society, it may be considered a waste of resources, especially when publicly funded.
+A formal assessment of the accumulated knowledge prior to research approval would reduce the waste of already limited resources caused by asking low priority questions.
+There is an urgent need for a change in research workflows so that pre-existing knowledge is better utilised in designing new research.*
 
 “Research waste” is a well established concept in medical research @glasziou2018research.
-Research is wasted when its outcomes cannot be used for the benefit of society @chalmers2009avoidable.
-Waste can occur at any of the four stages of research @chalmers2009avoidable; question setting; methods; accessibility; and reporting.
-In recent years there has been an increased focus in ecology on methods (e.g.
-@fraser2018questionable; @nilsen_bowler_linnell_2019) which have called for methodological improvements in individual studies.
-Open science is fast becoming mainstream in ecological publishing which is vastly improving accessibility @powers2019open.
-Evidence synthesis approaches such as systematic review and meta-analysis have become stalwarts of ecology and conservation research @gurevitch2018meta leading to improved reporting for the benefit of decision-makers.
-Despite these advances research waste still occurs in ecology and here we focus on the critical area of question setting where ecology and conservation can make large steps forward in research waste reduction.
+Research is wasted when its outcomes cannot be used for the benefit of society , for example because no new knowledge is gained or the knowledge gained cannot be applied in practice.
+Waste can occur at any of the four stages of research @chalmers2009avoidable; question setting; methods; accessibility; and reporting (Figure 1; Table 1).
+In medicine, research waste was estimated in 2009 to cost US$1bn @chalmers2009avoidable, with little improvement in the last decade1.
+There is little reason to suspect that the situation is substantially different in ecology and conservation, although there are no field-wide formal assessments of research waste.
+Emerging topics are however beginning to address some of the issues, for example through methodological improvements in individual studies (e.g.
+@fraser2018questionable; @nilsen_bowler_linnell_2019) as well as a focus on open science and evidence synthesis, leading to improved accessibility and reporting  @powers2019open; @gurevitch2018meta.
+Here we focus on the critical area of question setting, where ecology and conservation can make large steps forward in research waste reduction.
 
-The question setting stage has two related areas where research waste can be reduced.
-Namely,  new studies do not take into account the body of evidence that have taken place prior to them and as a consequence low priority questions are being asked.
-A low priority question may be one that is not relevant to stakeholders (to remedy this stakeholders should be apart of the research commissioning process @chalmers2009avoidable), or be one that has already been sufficiently addressed in the literature so that we already know the outcome with high certainty.
-Any further studies could then be considered a waste of resources, because they will not add (or only marginally so) to our knowledge base.
+The question setting stage has two related areas where research waste can be reduced by taking into account the existing body of evidence.
+Firstly, new studies may ask low priority questions - those that are irrelevant to stakeholders.
+The remedy to this is relatively straightforward; stakeholders should be apart of the research commissioning process @chalmers2009avoidable.
+Secondly, if a topic has been sufficiently addressed in the existing literature we might already know the outcome with high certainty.
+Further studies that fail to leverage this existing knowledge are at high risk of wasting precious research resources.
 
@@
 Well over 150 field studies have been carried out that address this issue either directly or indirectly and seek to address the question of whether acoustic recorders can replace human observers in wildlife surveys.
-A meta-analysis in 2018 @darras2018comparing
- explored the pooled effect of these studies using a meta-analytic approach to estimate species richness of birds.
+A meta-analysis in 20189 explored the pooled effect of these studies using a meta-analytic approach to estimate species richness of birds.
 Based on the combined evidence from the included studies, they concluded that when human observers (using point counts) and sound recorders sample areas of equal size then there is no difference between estimates of bird species richness.
@@
 
-Taking the role as a research funder or researcher at the question setting stage, it would be nice if we could know if we need another study quantifying the difference between acoustic recorders and human observers for bird survey point counts to make this conclusion?
+Taking the role as a research funder or researcher at the question setting stage, it would be valuable if we could know if we need another study quantifying the difference between acoustic recorders and human observers for bird survey point counts to make this conclusion?
 This is exactly the situation where the cumulative meta-analysis approach is instrumental.
@@
 We extracted the data and R code from @darras2018comparing to recreate their analysis.
-Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” @metafor
-package in R.
+Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” @metafor package in R.
 The cumulative meta-analysis was ordered by publication year and plotted using the “forest” function.
@@
 This approach (known as “trial sequential analysis”) is well developed in medicine @wetterslev2008trial.
-Plots were produced using ggplot2 @ggplot2.
 
-The effect size (i.e the magnitude of the difference between intervention and control) of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0 in each study (Figure 1).
+The effect size (i.e the magnitude of the difference between intervention and control) of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0 in each study (Figure 3).
 This means that there was no clear difference between acoustic recorders and human observers on bird point counts.
@@
 One option is to use cumulative meta-analysis.
-The approach demonstrated here is well established in the medical literature.
-It is well known and tested.
-Therefore, it should not be challenging to integrate in to conservation and applied ecology workflows.
+The approach demonstrated here is well known and tested in the medical literature.
+Therefore, it should not be challenging to integrate into conservation and applied ecology workflows.
+Cumulative meta-analysis has already been used in our field to assess time-lag bias12 but is not commonly used in the way we have shown here.
 Ideally, funding decisions will take into account the available evidence before giving grants to well studied topics.
+There may of course be a time lag between identifying that we have sufficient information, and not conducting further research on a topic, due to timescales of publications of original research as well as associated meta-analyses.
 
 More frequent application of systematic reviews combined with meta-analysis in general, and cumulative meta-analysis in particular, would be one important remedy to reduce research waste in ecology and conservation.
-However, there are at least two important caveats that need to be highlighted.
-First, is that currently, due to heterogeneity in reporting as well as the drive for novelty in publications that meta-analysis is challenging in applied ecology.
+However, there are at least two important caveats.
+First, due to heterogeneity in reporting as well as the drive for novelty in publications, meta-analysis is currently challenging in applied ecology.
 There might not be sufficient good quality research to quantify the cumulative effect of even some apparently well studied phenomena.
@@
 
-It is also important to remember that research funding might not only be about answering a research question and could be more targeted to the development of researchers skills (PhD students, etc.) or to engaging with stakeholders.
+Second, it is important to remember that research funding might not only be about answering a research question and could be more targeted to the development of researchers’ skills (PhD students, etc.) or to engaging with stakeholders.
 Funding for training, skill development and outreach activities should be given appropriate weight in any funding prioritisation.
@@
 Research waste can be reduced and it is the responsibility of funders as well as individual researchers to do so.
-We agree with the statement targeted at medicine 25 years ago that “We need less research, better research, and research done for the right reasons” @altman1994scandal.
-Without a change in focus ecology and conservation funding will continue to be wasted which will be detrimental to our efforts to provide solutions to global challenges.
+We agree with the statement targeted at medicine 25 years ago that “We need...better research, and research done for the right reasons” @altman1994scandal.
+Without a change in focus ecology and conservation funding will continue to be wasted which will be detrimental to our efforts to provide solutions to global societal challenges.
 
@@
 
-```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
-(2018) on the difference between human observers and acoustic recorders in terms of species richness.\\label{forest}"}
-### Add cumulative meta-analysis
-res <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=meta1, slab=paste(meta1$first_author, meta1$year, sep=", "))
-cres<-cumul(res, order=order(meta1$year))
-forest(cres, col="blue")
+```{r flowc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "The production of research flows through four stages all of which can lead to research waste [@fraser2018questionable].
+Ecology and conservation have begun to reduce waste by focusing on methodological improvements and open science.
+Evidence synthesis (including reporting to decision makers) can be considered another stage of research production.
+The failure to make use of the available evidence in the design of new studies is another avenue to research waste.
+The use of cumulative meta-analysis is one way in which this type of research waste can be reduced.\\label{flow}"}
+library(DiagrammeR)
+
+grViz("
+digraph boxes_and_circles {
+
+  # a 'graph' statement
+  graph [overlap = true, fontsize = 10]
+
+  # several 'node' statements
+  node [shape = box,
+        fontname = Helvetica]
+      tab1 [label = '@@1',color = blue]
+      tab2 [label = '@@2',color = blue]
+      tab3 [label = '@@3',color = blue]
+      tab4 [label = '@@4',color = blue]
+      tab5 [label = '@@5',color = red]
+  node [shape = circle,
+        fixedsize = true,
+        width = 2, color=green] // sets as circles
+    tab6 [label = '@@6']
+    # several 'edge' statements
+ tab1->tab2->tab3->tab4
+ tab1->tab5
+  tab2->tab5
+   tab3->tab5
+    tab4->tab5
+    tab4->tab6 
+    tab6->tab1 [style=dotted]
+        tab6->tab2 [style=dotted]
+        tab6->tab5 [style=dashed]
+
+}
+      [1]: 'Questions relevant to stakeholders'
+      [2]: 'Appropriate design & methods'
+      [3]: 'Accessible full publication'
+      [4]: 'Unbiased reporting'
+      [5]: 'Research waste'
+      [6]: 'Evidence synthesis'
+", height=1000)
+
 ```
@@
 The red dashed lines indicate the significance threshold (between -1.96 and 1.96).
-When the z-curve crosses the red line then there is sufficient evidence and no new trials are required.
+When the z-curve crosses the red line then there is sufficient evidence and no new trials are required [@ggplot2].
 \\label{seq}"}
-
+res <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=meta1, slab=paste(meta1$first_author, meta1$year, sep=", "))
+cres<-cumul(res, order=order(meta1$year))
 cumzval<-cumsum(cres$zval)
@@
 ```
+
+```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
+(2018) on the difference between human observers and acoustic recorders in terms of species richness.\\label{forest}"}
+### Add cumulative meta-analysis
+
+forest(cres, col="blue")
+```

### 2019-11-25 Manuscript/MS.Rmd
update 25/11/2019

@@
       tab4 [label = '@@4',color = blue]
-      tab5 [label = '@@5',color = red]
+      tab6 [label = '@@6',color = blue]
+      tab8 [label = '@@8',color = red]
+      tab9 [label = '@@9',color = red]
+      tab10[label = '@@10',color = red]
+      tab11[label = '@@11',color = red]
+
   node [shape = circle,
         fixedsize = true,
-        width = 2, color=green] // sets as circles
-    tab6 [label = '@@6']
+        width = 2, color=blue] // sets as circles
+    tab7 [label = '@@7']
+
+    node [shape = circle,
+        fixedsize = true,
+        width = 2, color=red] // sets as circles
+    tab5 [label = '@@5']
+
     # several 'edge' statements
- tab1->tab2->tab3->tab4
- tab1->tab5
-  tab2->tab5
-   tab3->tab5
-    tab4->tab5
-    tab4->tab6 
-    tab6->tab1 [style=dotted]
-        tab6->tab2 [style=dotted]
-        tab6->tab5 [style=dashed]
+ tab1->tab2->tab3->tab4->tab6->tab7 [color=blue]
+ tab1->tab8 [color=red,style=dashed]
+  tab2->tab9 [color=red,style=dashed]
+   tab3->tab10 [color=red,style=dashed]
+    tab4->tab11 [color=red,style=dashed]
+    tab8->tab5[color=red,style=dashed]
+    tab9->tab5[color=red,style=dashed]
+    tab10->tab5[color=red,style=dashed]
+    tab11->tab5[color=red,style=dashed]
+    tab6->tab1 [style=dashed]
+        tab6->tab2 [style=dashed]
+    tab6->tab5 [color=red, style=dashed]
 
@@
       [4]: 'Unbiased reporting'
-      [5]: 'Research waste'
+      [5]: 'Wasteful research'
       [6]: 'Evidence synthesis'
+      [7]: 'Utilisable research' 
+      [8]: 'Previous knowledge not accounted for'
+      [9]: 'Poor design; \\nquestionable research practices'
+      [10]: 'Publications not available to practitioners \\n& decision makers'
+      [11]: 'Lack of transparency'
   ", height=1000)
@@
 cum_df %>% 
-  ggplot(aes(cum_sample,z.score, colour=as.factor(Year)))+
+  ggplot(aes(cum_sample,z.score))+
   scale_fill_npg()+
   labs(y="Cumulative z score", x="Cumulative sample size")+
-  geom_point(size=6)+
+  geom_point(size=3, colour="blue")+
   geom_line(colour="blue", linetype="dashed", size=1)+

### 2019-11-25 Flowchart.R

@@ -63,3 +63,63 @@ digraph boxes_and_circles {
       [5]: 'Research waste'
       [6]: 'Evidence synthesis'
 ")
+
+
+grViz("
+digraph boxes_and_circles {
+
+  # a 'graph' statement
+  graph [overlap = true, fontsize = 10]
+
+  # several 'node' statements
+  node [shape = box,
+        fontname = Helvetica]
+      tab1 [label = '@@1',color = blue]
+      tab2 [label = '@@2',color = blue]
+      tab3 [label = '@@3',color = blue]
+      tab4 [label = '@@4',color = blue]
+      tab6 [label = '@@6',color = blue]
+      tab8 [label = '@@8',color = red]
+      tab9 [label = '@@9',color = red]
+      tab10[label = '@@10',color = red]
+      tab11[label = '@@11',color = red]
+      
+  node [shape = circle,
+        fixedsize = true,
+        width = 2, color=blue] // sets as circles
+    tab7 [label = '@@7']
+    
+    node [shape = circle,
+        fixedsize = true,
+        width = 2, color=red] // sets as circles
+    tab5 [label = '@@5']
+  
+    
+    # several 'edge' statements
+ tab1->tab2->tab3->tab4->tab6->tab7 [color=blue]
+ tab1->tab8 [color=red,style=dashed]
+  tab2->tab9 [color=red,style=dashed]
+   tab3->tab10 [color=red,style=dashed]
+    tab4->tab11 [color=red,style=dashed]
+    tab8->tab5[color=red,style=dashed]
+    tab9->tab5[color=red,style=dashed]
+    tab10->tab5[color=red,style=dashed]
+    tab11->tab5[color=red,style=dashed]
+    tab6->tab1 [style=dashed]
+        tab6->tab2 [style=dashed]
+    tab6->tab5 [color=red, style=dashed]
+        
+ }
+      [1]: 'Questions relevant to stakeholders'
+      [2]: 'Appropriate design & methods'
+      [3]: 'Accessible full publication'
+      [4]: 'Unbiased reporting'
+      [5]: 'Wasteful research'
+      [6]: 'Evidence synthesis'
+      [7]: 'Utilisable research' 
+      [8]: 'Previous knowledge not accounted for'
+      [9]: 'Poor design; \\nquestionable research practices'
+      [10]: 'Publications not available to practitioners \\n& decision makers'
+      [11]: 'Lack of transparency'
+  ", height=1000)
+

### 2019-11-27 Manuscript/MS.Rmd
27.11.2019

@@
       [2]: 'Appropriate design & methods'
-      [3]: 'Accessible full publication'
-      [4]: 'Unbiased reporting'
+      [3]: 'Unbiased reporting'
+      [4]: 'Accessible full publication'
       [5]: 'Wasteful research'
@@
       [9]: 'Poor design; \\nquestionable research practices'
-      [10]: 'Publications not available to practitioners \\n& decision makers'
-      [11]: 'Lack of transparency'
+      [10]: 'Lack of transparency'
+      [11]: 'Publications not available to practitioners \\n& decision makers'
   ", height=1000)
@@
   theme(text=element_text(size=16,  family="serif"))
+#dev.off()
 ```

### 2019-11-27 Flowchart.R

@@ -65,6 +65,7 @@ digraph boxes_and_circles {
 ")
 
 
+
 grViz("
 digraph boxes_and_circles {
 
@@ -112,14 +113,13 @@ digraph boxes_and_circles {
  }
       [1]: 'Questions relevant to stakeholders'
       [2]: 'Appropriate design & methods'
-      [3]: 'Accessible full publication'
-      [4]: 'Unbiased reporting'
+      [3]: 'Unbiased reporting'
+      [4]: 'Accessible full publication'
       [5]: 'Wasteful research'
       [6]: 'Evidence synthesis'
       [7]: 'Utilisable research' 
       [8]: 'Previous knowledge not accounted for'
       [9]: 'Poor design; \\nquestionable research practices'
-      [10]: 'Publications not available to practitioners \\n& decision makers'
-      [11]: 'Lack of transparency'
+      [10]: 'Lack of transparency'
+      [11]: 'Publications not available to practitioners \\n& decision makers'
   ", height=1000)
-

### 2019-11-27 Rcode.R

@@ -92,11 +92,19 @@ RMA_alpha_identical=rma.mv(yi=alpha_ROM_identical,V=alpha_variance_ROM_identical
 ### Add cumulative meta-analysis
 res <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=meta1, slab=paste(meta1$first_author, meta1$year, sep=", "))
 cres<-cumul(res, order=order(meta1$year))
+
+meta1$IF<-c(2.27,0.967,0.967,0.967,1.29,0.41,0.967,0.967,0.967,0,0.538,2.055,2.53,2.53,0.622,1.92,1.895,0.94,1.895,0,0.967,1.75,6.36,0,0.41,0,0.41)
+cres2<-cumul(res, order=order(meta1$IF))
+
+#plot.cumul.rma(cres)
+
+
 #Figure1
 png("Figures/Figure1.png", width=850, height =400)
 forest(cres, col="blue")
 dev.off()
 
+forest(cres2, col="blue")
 #Cumulative data frame
 cumzval<-cumsum(cres$zval)
 cum_sample<-cumsum(meta1$alpha_point_n_identical)  

### 2019-12-02 Manuscript/MS.Rmd
2.12.2019

@@
 ---
-title: "Maximising the leverage of existing knowledge could reduce research waste in applied ecology and conservation"
+title: "Evidence accumulation for tackling research waste"
 csl: nature.csl
@@
 
-*Where research does not lead to benefits to society, it may be considered a waste of resources, especially when publicly funded.
-A formal assessment of the accumulated knowledge prior to research approval would reduce the waste of already limited resources caused by asking low priority questions.
-There is an urgent need for a change in research workflows so that pre-existing knowledge is better utilised in designing new research.*
+*There is an urgent need for a change in research workflows so that pre-existing knowledge is better utilised in designing new research.
+A formal assessment of the accumulated knowledge prior to research approval would reduce the waste of already limited resources caused by asking low priority questions.*
 
-“Research waste” is a well established concept in medical research @glasziou2018research.
-Research is wasted when its outcomes cannot be used for the benefit of society , for example because no new knowledge is gained or the knowledge gained cannot be applied in practice.
-Waste can occur at any of the four stages of research @chalmers2009avoidable; question setting; methods; accessibility; and reporting (Figure 1; Table 1).
-In medicine, research waste was estimated in 2009 to cost US$1bn @chalmers2009avoidable, with little improvement in the last decade1.
-There is little reason to suspect that the situation is substantially different in ecology and conservation, although there are no field-wide formal assessments of research waste.
-Emerging topics are however beginning to address some of the issues, for example through methodological improvements in individual studies (e.g.
-@fraser2018questionable; @nilsen_bowler_linnell_2019) as well as a focus on open science and evidence synthesis, leading to improved accessibility and reporting  @powers2019open; @gurevitch2018meta.
-Here we focus on the critical area of question setting, where ecology and conservation can make large steps forward in research waste reduction.
+“Research waste” is a well-established concept in medical research[@glasziou2018research].
+Research is wasted when its outcomes cannot be used for the benefit of society [@chalmers2009avoidable], for example because no new knowledge is gained or the knowledge gained cannot be applied.
+Waste can occur at any of the four stages of the research process [@chalmers2009avoidable]; question setting; methods; accessibility; and reporting (Table 1).
+In medicine, global research waste was estimated in 2009 to cost US $85 bn [@chalmers2009avoidable], with few signs of improvement in the last decade [@glasziou2018research].
+There is little reason to believe that the situation is substantially different in ecology and conservation, although there are no field-wide formal assessments of research waste.
 
-The question setting stage has two related areas where research waste can be reduced by taking into account the existing body of evidence.
-Firstly, new studies may ask low priority questions - those that are irrelevant to stakeholders.
-The remedy to this is relatively straightforward; stakeholders should be apart of the research commissioning process @chalmers2009avoidable.
-Secondly, if a topic has been sufficiently addressed in the existing literature we might already know the outcome with high certainty.
-Further studies that fail to leverage this existing knowledge are at high risk of wasting precious research resources.
+Emerging topics are beginning to address some of the factors that result in wasted research efforts (Table 1).
+In particular, there is increased focus on methodological improvements in individual studies (e.g.
+[@fraser2018questionable], [@nilsen_bowler_linnell_2019]),  and on open science leading to improved accessibility and reporting [@gurevitch2018meta], [@powers2019open].
+Less formal effort is devoted to the question setting stage.
+Here we suggest that “Evidence Synthesis” should be considered an additional stage of research (Table1, Figure S1).
+Evidence synthesis methods close the research process into a loop, and will have additional benefits in terms of reducing research waste at the question setting stage.
 
-In medicine, one way in which this particular type of research waste has been identified is by conducting cumulative meta-analysis based on effect sizes from previous studies.
-A cumulative meta-analysis typically describes the accumulation of evidence (e.g., about the effectiveness of an intervention) across time, and available estimates are added to the analysis in chronological order  @lau1992cumulative.
-Using cumulative meta-analysis, a researcher, funding agency or decision maker can identify the point at which there is sufficient evidence to be confident that a reported effect is true.
-At this stage new trials are no longer required to predict the outcome with satisfactory certainty.
+# Reducing waste in question setting
 
-As an example of the approach applied to an applied ecology situation we can look at the potential of autonomous acoustic recorders to replace human observers in wildlife sampling and monitoring, which now has a long history in the ecological literature (starting from at least @ralph1995monitoring).
-Technological advances over the last two decades have allowed this potential to be explored fully.
-Well over 150 field studies have been carried out that address this issue either directly or indirectly and seek to address the question of whether acoustic recorders can replace human observers in wildlife surveys.
-A meta-analysis in 20189 explored the pooled effect of these studies using a meta-analytic approach to estimate species richness of birds.
+There are two related areas where research waste can be reduced by taking into account the existing body of evidence by applying evidence synthesis methods.
+
+## Low priority questions
+
+New studies may ask low priority questions - those that are irrelevant to stakeholders.
+The remedy to this is to include stakeholders in the research commissioning process[@chalmers2009avoidable].
+Evidence synthesis, or horizon scanning for novel problems, should be used to provide evidence to practitioners, researchers and other stakeholders so that they can identify research gaps that are important to them and to develop future questions [@gold2013prioritizing].
+
+## The answer is already known with certainty
+
+If a topic has been sufficiently addressed in the existing literature we might already know the outcome with high certainty.
+Further studies that fail to leverage this existing knowledge are at high risk of wasting limited research resources.
+There are a variety of tools available for research-funders and researchers to assess the state of knowledge on the topic of interest.
+For example, systematic maps (also known as Evidence gap maps or Evidence maps), were designed to give an overview of the available evidence on a broad topic [@saran2018evidence].
+They can highlight where there is enough available evidence for a systematic review or where primary research is required (i.e.
+there is a lack of evidence).
+Systematic reviews can be used to synthesise knowledge about a narrow topic such as the evidence for the effectiveness of an intervention and can provide a statistical summary of the pooled effect size.
+The statistical combination of numerical data extracted from the evidence base during the process of a systematic review is known as meta-analysis.
+Meta-analysis is commonly used in conservation and ecology6 providing an understanding of the magnitude of the known effect of an intervention across individual studies.
+These results can then be used to identify what a new research project can add to the current evidence base.
+
+# Identifying research waste with cumulative meta-analysis
+In medicine, one additional tool used to quantify research waste is cumulative meta-analysis.
+A cumulative meta-analysis typically describes the accumulation of evidence (e.g., about the effectiveness of an intervention) across time, and available estimates are added to the analysis in chronological order [@lau1992cumulative].
+Using cumulative meta-analysis, a researcher, funding agency or decision maker can identify if there is sufficient evidence to be confident that a reported effect is true.
+At this stage new trials are no longer required to predict the outcome with satisfactory certainty and hence future research waste will be avoided.
+
+## An applied example
+As an example of the approach within an applied ecology situation, we consider to what extent autonomous acoustic recorders can replace human observers in wildlife sampling and monitoring when the focus is on estimating species richness, which now has a long history in the ecological literature [@darras2018comparing].
+Technological advances over the last two decades have allowed this potential to be explored fully, and well over 150 field studies have sought to answer this question.
+A meta-analysis in 2018[@darras2018comparing] explored the pooled effect of these studies using a meta-analytic approach to estimate species richness of birds.
 Based on the combined evidence from the included studies, they concluded that when human observers (using point counts) and sound recorders sample areas of equal size then there is no difference between estimates of bird species richness.
-When properly conducted (see specific advice in @darras2018comparing), it can be inferred that sound recorders can be used to monitor aspects of biodiversity as efficiently as human observers.
-Twenty-eight primary studies published between 2000 and 2017 were included in their meta-analysis.
+When properly conducted (see specific advice in [@darras2018comparing]), it can be inferred that sound recorders can be used to monitor aspects of biodiversity as efficiently as human observers.
+Twenty-eight primary studies published between 2000 and 2017 were included in the meta-analysis.
+Taking the role as a research funder or researcher at the question setting stage, we can utilise cumulative meta-analysis to determine if we need another study quantifying the difference between acoustic recorders and human observers for bird survey point counts.
+We adapted the analysis of [@darras2018comparing] to demonstrate the use of cumulative meta-analysis (see supplementary materials).
+The effect size (i.e the magnitude of the difference between intervention and control) of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0.07 since 2015 (Figure 1).
+This means that there was no clear difference between acoustic recorders and human observers on bird point counts.
+It would therefore be wasteful to fund yet another study that addressed this specific question.
+To reduce research waste we need to be able to first identify it.
+One option is to use cumulative meta-analysis.
+The approach demonstrated here is well known and tested in the medical literature and should not be challenging to integrate into conservation and applied ecology workflows.
+Cumulative meta-analysis has already been used in our field to assess time-lag bias [@leimu2004cumulative] but is not commonly used in the way we have shown here.
 
-Taking the role as a research funder or researcher at the question setting stage, it would be valuable if we could know if we need another study quantifying the difference between acoustic recorders and human observers for bird survey point counts to make this conclusion?
-This is exactly the situation where the cumulative meta-analysis approach is instrumental.
+## Caveats
+There are several important caveats that need to be addressed.
+The heterogeneity in reporting and the drive for novelty in publications means that meta-analysis is currently challenging in applied ecology.
+There might not be sufficient good quality research to quantify the cumulative effect of even some apparently well studied phenomena.
+Researchers are best placed to add to the evidence base by ensuring that they use of comparable measures of outcomes rather than novel ones.
+In addition, publication bias, where the direction of statistical significance of the outcome influences the decision to publish the result, might bias the evidence base available.
+This is a major caveat for all evidence synthesis approaches, but one which can be identified.
+With cumulative meta-analysis one can identify publication bias11 by accumulating the effect sizes in order of journal impact factor for example.
+Although this method makes it possible to detect publication bias it will not not solve the underlying problem, and researchers should endeavour to reduce publication bias by following open science (Table 1).
+There may of course be a time lag between identifying that we have sufficient information, and not conducting further research on a topic, due to timescales of publications of original research as well as associated meta-analyses.
+To address these and other issues around evidence synthesis we have been exploring a systems modelling approach to combine empirical evidence from systematic reviews and meta-analysis with expert opinion which allows key areas of uncertainty in a topic to be identified and prioritised for research focus (e.g.
+[@carrick2018planting]).
 
-We extracted the data and R code from @darras2018comparing to recreate their analysis.
-Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” @metafor package in R.
-The cumulative meta-analysis was ordered by publication year and plotted using the “forest” function.
-To assess the point at which there is sufficient evidence and no further investigations are required we plotted the z-curve in relation to the cumulative sample size.
-The thresholds for significance was a z value of 1.96 or -1.96.
-When the z curve crosses this threshold then the level of evidence is considered sufficient.
-This approach (known as “trial sequential analysis”) is well developed in medicine @wetterslev2008trial.
+## Outlook
+Research waste can be reduced and it is the responsibility of funders as well as individual researchers to do so.
+We agree with the statement targeted at medicine 25 years ago that “We need...better research, and research done for the right reasons” [@altman1994scandal].
+Without a change in focus ecology and conservation funding will continue to be wasted which will be detrimental to our efforts to provide solutions to global societal challenges.
 
-The effect size (i.e the magnitude of the difference between intervention and control) of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0 in each study (Figure 3).
-This means that there was no clear difference between acoustic recorders and human observers on bird point counts.
-Trial sequential analysis (Figure 2) shows that the significance threshold (between -1.96 and 1.96) was crossed in 2015 favouring automatic recorders over human observers in bird point surveys.
-This suggests that studies undertaken post 2015 were a “waste” of research resources.
+# Supplementary Information
 
-To reduce research waste we need to be able to first quantify it.
-One option is to use cumulative meta-analysis.
-The approach demonstrated here is well known and tested in the medical literature.
-Therefore, it should not be challenging to integrate into conservation and applied ecology workflows.
-Cumulative meta-analysis has already been used in our field to assess time-lag bias12 but is not commonly used in the way we have shown here.
-Ideally, funding decisions will take into account the available evidence before giving grants to well studied topics.
-There may of course be a time lag between identifying that we have sufficient information, and not conducting further research on a topic, due to timescales of publications of original research as well as associated meta-analyses.
-
-More frequent application of systematic reviews combined with meta-analysis in general, and cumulative meta-analysis in particular, would be one important remedy to reduce research waste in ecology and conservation.
-However, there are at least two important caveats.
-First, due to heterogeneity in reporting as well as the drive for novelty in publications, meta-analysis is currently challenging in applied ecology.
-There might not be sufficient good quality research to quantify the cumulative effect of even some apparently well studied phenomena.
-In cases such as this one might consider using a systems modelling approach to identify key areas of uncertainty in a topic that could be prioritised for research focus.
-
-Second, it is important to remember that research funding might not only be about answering a research question and could be more targeted to the development of researchers’ skills (PhD students, etc.) or to engaging with stakeholders.
-Funding for training, skill development and outreach activities should be given appropriate weight in any funding prioritisation.
-
-Research waste can be reduced and it is the responsibility of funders as well as individual researchers to do so.
-We agree with the statement targeted at medicine 25 years ago that “We need...better research, and research done for the right reasons” @altman1994scandal.
-Without a change in focus ecology and conservation funding will continue to be wasted which will be detrimental to our efforts to provide solutions to global societal challenges.
+## Methods
+We extracted the data and R code from the supplementary information in [@darras2018comparing] to recreate their analysis.
+As such we are dependant on the accurate extraction of data from the primary studies by the original authors.
+We intended our re-analysis of their data to be an example of the cumulative meta-analysis approach rather than to make explicit recommendations about the use of acoustic recorders for avian surveys.
+Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” [@metafor] package in R.
+The cumulative meta-analysis was ordered by publication year and plotted using the ggplot2 package in R [@ggplot2] .
+Where a single study provided more than one estimate of effect the order in which the estimates were accumulated was the same as the order presented by [@darras2018comparing] and treated as subsequent trials.
+Changing the order of that these particular studies were accumulated made no difference to the stability of the estimates over time (see Figure 3).
+The original code, the original data, our additional code for running the analysis and plotting can be found at https://github.com/DrMattG/Research_waste.
 
@@
 library(gganimate)
+library(ggthemes)
+library(bookdown)
+
 #Code from the original paper (Darras et al. 2018)
@@
 meta1[,publication:=paste(first_author,year)]
+meta1<-meta1 %>% 
+  mutate("year"=gsub("a"," ", year)) %>% 
+  mutate("year"=gsub("b"," ", year))
 
@@
                            ,data=meta1,random=list(~1|publication))
+forest(RMA_alpha_identical)
 
@@
 
-```{r flowc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "The production of research flows through four stages all of which can lead to research waste [@fraser2018questionable].
+```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
+(2018) on the difference between human observers and acoustic recorders in terms of species richness.\\label{forest}"}
+### Add cumulative meta-analysis
+res <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=meta1, slab=paste(meta1$first_author, meta1$year, sep=", "))
+cres<-cumul(res, order=order(meta1$year))
+#forest(cres, col="blue")
+df<-data.frame("cite"=cres$slab, "est"=cres$estimate, "se"=cres$se)
+
+p<-df %>% 
+  rowid_to_column() %>% 
+  mutate("Up"=(1.96*se)+est) %>% 
+  mutate("Low"=(-1.96*se)+est) %>% 
+  ggplot(aes(y=reorder(cite,-rowid), x=est, xmin=Low, xmax=Up))+
+  geom_point(color = 'blue', size=3)+
+  labs(x="Log ratio of means", y="")+
+  geom_vline(xintercept = 0, linetype = "dashed", colour="green", size=1)+
+  geom_errorbarh(height=.1, colour="blue", size=1)+
+  ggthemes::theme_base()+
+  theme(axis.line = element_line(colour = 'black', size = 1), 
+        axis.ticks = element_line(colour = "black", size = 1))
+p
+
+```
+
+```{r flowc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "The production of research flows through five stages (blue lines) all of which can lead to research waste [@@chalmers2009avoidable] (red dashed lines).
 Ecology and conservation have begun to reduce waste by focusing on methodological improvements and open science.
-Evidence synthesis (including reporting to decision makers) can be considered another stage of research production.
-The failure to make use of the available evidence in the design of new studies is another avenue to research waste.
-The use of cumulative meta-analysis is one way in which this type of research waste can be reduced.\\label{flow}"}
+Evidence synthesis (including reporting to decision makers) can contribute to the reduction in research waste by influencing question setting and appropriate methods and design (black dashed lines).\\label{flow}"}
 library(DiagrammeR)
@@
 
-```{r Add seq, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Trial sequential analysis.
-The red dashed lines indicate the significance threshold (between -1.96 and 1.96).
-When the z-curve crosses the red line then there is sufficient evidence and no new trials are required [@ggplot2].
-\\label{seq}"}
-res <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=meta1, slab=paste(meta1$first_author, meta1$year, sep=", "))
-cres<-cumul(res, order=order(meta1$year))
-cumzval<-cumsum(cres$zval)
-cum_sample<-cumsum(meta1$alpha_point_n_identical)  
-# cum_sample<-c(136,              186,
-#               190,              238,
-#               363,              419,
-#               455,              516,
-#               577,              647,
-#               660,              720,
-#               780,              878,
-#               894,              898,
-#               922,              930,
-#               1103,              1113,
-#               1172,              1633,
-#               1666,              1710,
-#               1752,              1784,
-#               1814)
+```{r Randomise order of the estimates in the same study,echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
+(2018) on the difference between human observers and acoustic recorders in terms of species richness.
+Estimates from the same study inputted in random order\\label{forestX}"}
+set.seed(8955)
+metax<-meta1
+metax$order<-order(metax$year, runif(length(metax$year))) 
+res2 <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=metax, slab=paste(metax$first_author, metax$year, sep=","))
+cres2<-cumul(res2, order=order(metax$order))
+par(mfrow=c(1,2))
+forest(cres2)
+forest(cres)
 
-Year=meta1$year
-
-Year=gsub("a","",Year)
-Year=gsub("b","",Year)
-Year<-as.integer(Year)
-Year<-Year[order(Year)]
-# Year=c(2000,2002,2002,2002,2004,2009,2009,2009,2012,2012,2012,2012,2012,
-# 2014,2015,2015,2015,2015,2015,2016,2016,2016,2016,2017,2017,2017,2018)
-library(ggrepel)
-cum_df<-data.frame("z-score"=cumzval, "Cumulative n"=cum_sample, "Year"=Year)
-
-cum_df2<-cum_df[c(1,18,27),]
-#png("TSA.png", width=850, height =400)
-cum_df %>% 
-  ggplot(aes(cum_sample,z.score))+
-  scale_fill_npg()+
-  labs(y="Cumulative z score", x="Cumulative sample size")+
-  geom_point(size=3, colour="blue")+
-  geom_line(colour="blue", linetype="dashed", size=1)+
-  ylim(c(-20,20))+
-  xlim(c(0,2000))+
-
-  #annotate("text", x = 1, y = 10, label = sprintf('\u2b61'),size=30) +
-  #annotate("text", x = 3000, y = 10, label ="Favours automatic recorder",size=4) +
-  #annotate("text", x = 1, y = -10, label = sprintf('\u2b63'),size=30) +
-  #annotate("text", x = 3000, y = -10, label ="Favours human recorder",size=4) +
-  geom_text(data=cum_df2, aes(Cumulative.n,z.score,label=as.factor(Year)),hjust=1.3, vjust=1, colour="black")+
-
-  #geom_text(aes(label=as.factor(Year)),hjust=0.5, vjust=0)+
-  geom_hline(yintercept = 2, colour="red",linetype="dotted", size=1.6)+
-  geom_hline(yintercept = -2, colour="red",linetype="dotted", size=1.6)+
-  theme_classic() +
-  theme(axis.line = element_line(size = 1.6, color = rgb(0,0,0,max=255)))+
-  theme(legend.position = "none")+
-  theme(text=element_text(size=16,  family="serif"))
-#dev.off()
 ```
 
-```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
-(2018) on the difference between human observers and acoustic recorders in terms of species richness.\\label{forest}"}
-### Add cumulative meta-analysis
+```{r, echo=FALSE}
+# df<-data.frame("Research Stage"=c("Questions relevant to stakeholders","Questions relevant to stakeholders","Appropriate design and methods","Appropriate design and methods",
+#                                   "Appropriate design and methods","Appropriate design and methods","Unbiased reporting","Unbiased reporting","Unbiased reporting","Unbiased reporting",
+#                                   "Unbiased reporting","Accessible full publication","Evidence synthesis"),
+#                "Examples of potential for research waste"=c("Irrelevant questions asked","Previous knowledge not properly taken into account","Study poorly designed, under-powered (or over-powered.
+etc.)",
+#                                                             "Using inappropriate statistical tools (including overfitting etc.)","Questionable research practices3 lead to poor quality research","Questionable research practices3 lead to poor quality research",
+#                                                             "Lack of open data","Hypothesising after the results are known","p-hacking","File-drawer syndrome (only some studies are published)","Incomplete reporting, making evidence synthesis difficult or impossible",
+#                                                             "Publications not available to practitioners and decision makers","Research not designed or presented in the context of the existing knowledge"),
+#                "Where ecology and conservation can reduce waste"=c("Co-development of research questions with stakeholders and using appropriate methodology such as Delphi exercises to avoid issues such as group think or not including the right group of experts or stakeholders",
+#                "Make use of evidence synthesis methods (e.g.
+cumulative meta-analysis, systematic mapping, systematic reviews, meta-analysis) to identify questions that are not satisfactorily answered",
+#                "Use simulations or power-analysis prior to undertaking data collection.
+Predefine effect size of interest with stakeholders (i.e.
+do not rely on rules of thumb for “statistical significance”)",
+#                "Better training of early-career researchers in methods. Open code and data to ensure reproducibility of methods",
+#                "Open science (open methods and data, reproducible methods, sharing code, etc.)",
+#                "Better training of early-career researchers in methods of open science and evidence synthesis",
+#                "Open science (open methods and data, reproducible methods, sharing code, etc.)",
+#                "Pre-registration of hypotheses",
+#                "Open science (open methods and data, reproducible methods, sharing code, etc.)",
+#                "Pre-registration of hypotheses and methods. Open publishing (including preprints)",
+#                "Increasing knowledge of researchers and peer reviewers on what is essential to report, and changing journal guidelines where necessary to ensure all relevant information is reported",
+#                "Open access publishing, including making resources available to researchers to be able to publish open access","Using systematic reviews, systematic maps, meta-analysis, etc.
+to shape research priorities.
+Where good quality evidence is available these should be synthesised providing evidence to relevant stakeholders.
+Research gaps should be the focus of primary studies."
+#                ))
+# # # kable(df, "latex", booktabs = T, align = "c") %>% column_spec(1, bold=T) %>% collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle")
+# library(kableExtra)
+# #knitr::kable(df)
+# df %>% kable("latex") %>% as_image()
+library(huxtable)
+ht <- hux(
+  Employee = c('John Smith', 'Jane Doe', 'David Hugh-Jones'),
+  Salary = c(50000, 50000, 40000),
+  add_colnames = TRUE
+)
 
-forest(cres, col="blue")
+bold(ht)[1,]           <- TRUE
+bottom_border(ht)[1,]  <- 2
+align(ht)[,2]          <- 'right'
+right_padding(ht)      <- 10
+left_padding(ht)       <- 10
+width(ht)              <- 0.35
+number_format(ht)      <- 2
+
+ht
+
 ```

### 2019-12-02 Manuscript/MS.tex
2.12.2019

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Evidence accumulation for tackling research waste},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Evidence accumulation for tackling research waste}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{array}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{calc}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{wrapfig}

\begin{document}
\maketitle

Matthew J. Grainger\textsuperscript{1}*, Frederike C.
Bolam\textsuperscript{2}, Gavin B. Stewart\textsuperscript{2}, Erlend B.
Nilsen\textsuperscript{1}

\textsuperscript{1}Norwegian Institute for Nature Research, P.O. Box
5685 Torgarden, 7485, Trondheim, Norway

\textsuperscript{2}Modelling Evidence and Policy Research Group, School
of Natural and Environmental Sciences, Ridley Building 2, Claremont
Road, Newcastle upon Tyne NE1 7RU, UK

*corresponding author: matthew.grainger@nina.no

\emph{There is an urgent need for a change in research workflows so that
pre-existing knowledge is better utilised in designing new research. A
formal assessment of the accumulated knowledge prior to research
approval would reduce the waste of already limited resources caused by
asking low priority questions.}

``Research waste'' is a well-established concept in medical
research\textsuperscript{1}. Research is wasted when its outcomes cannot
be used for the benefit of society\textsuperscript{2}, for example
because no new knowledge is gained or the knowledge gained cannot be
applied. Waste can occur at any of the four stages of the research
process\textsuperscript{2}; question setting; methods; accessibility;
and reporting (Table 1). In medicine, global research waste was
estimated in 2009 to cost US \$85 bn\textsuperscript{2}, with few signs
of improvement in the last decade\textsuperscript{1}. There is little
reason to believe that the situation is substantially different in
ecology and conservation, although there are no field-wide formal
assessments of research waste.

Emerging topics are beginning to address some of the factors that result
in wasted research efforts (Table 1). In particular, there is increased
focus on methodological improvements in individual studies
(e.g.~\textsuperscript{3},\textsuperscript{4}), and on open science
leading to improved accessibility and
reporting\textsuperscript{5},\textsuperscript{6}. Less formal effort is
devoted to the question setting stage. Here we suggest that ``Evidence
Synthesis'' should be considered an additional stage of research
(Table1, Figure S1). Evidence synthesis methods close the research
process into a loop, and will have additional benefits in terms of
reducing research waste at the question setting stage.

\hypertarget{reducing-waste-in-question-setting}{%
\section{Reducing waste in question
setting}\label{reducing-waste-in-question-setting}}

There are two related areas where research waste can be reduced by
taking into account the existing body of evidence by applying evidence
synthesis methods.

\hypertarget{low-priority-questions}{%
\subsection{Low priority questions}\label{low-priority-questions}}

New studies may ask low priority questions - those that are irrelevant
to stakeholders. The remedy to this is to include stakeholders in the
research commissioning process\textsuperscript{2}. Evidence synthesis,
or horizon scanning for novel problems, should be used to provide
evidence to practitioners, researchers and other stakeholders so that
they can identify research gaps that are important to them and to
develop future questions\textsuperscript{7}.

\hypertarget{the-answer-is-already-known-with-certainty}{%
\subsection{The answer is already known with
certainty}\label{the-answer-is-already-known-with-certainty}}

If a topic has been sufficiently addressed in the existing literature we
might already know the outcome with high certainty. Further studies that
fail to leverage this existing knowledge are at high risk of wasting
limited research resources. There are a variety of tools available for
research-funders and researchers to assess the state of knowledge on the
topic of interest. For example, systematic maps (also known as Evidence
gap maps or Evidence maps), were designed to give an overview of the
available evidence on a broad topic\textsuperscript{8}. They can
highlight where there is enough available evidence for a systematic
review or where primary research is required (i.e.~there is a lack of
evidence). Systematic reviews can be used to synthesise knowledge about
a narrow topic such as the evidence for the effectiveness of an
intervention and can provide a statistical summary of the pooled effect
size. The statistical combination of numerical data extracted from the
evidence base during the process of a systematic review is known as
meta-analysis. Meta-analysis is commonly used in conservation and
ecology6 providing an understanding of the magnitude of the known effect
of an intervention across individual studies. These results can then be
used to identify what a new research project can add to the current
evidence base.

\hypertarget{identifying-research-waste-with-cumulative-meta-analysis}{%
\section{Identifying research waste with cumulative
meta-analysis}\label{identifying-research-waste-with-cumulative-meta-analysis}}

In medicine, one additional tool used to quantify research waste is
cumulative meta-analysis. A cumulative meta-analysis typically describes
the accumulation of evidence (e.g., about the effectiveness of an
intervention) across time, and available estimates are added to the
analysis in chronological order\textsuperscript{9}. Using cumulative
meta-analysis, a researcher, funding agency or decision maker can
identify if there is sufficient evidence to be confident that a reported
effect is true. At this stage new trials are no longer required to
predict the outcome with satisfactory certainty and hence future
research waste will be avoided.

\hypertarget{an-applied-example}{%
\subsection{An applied example}\label{an-applied-example}}

As an example of the approach within an applied ecology situation, we
consider to what extent autonomous acoustic recorders can replace human
observers in wildlife sampling and monitoring when the focus is on
estimating species richness, which now has a long history in the
ecological literature\textsuperscript{10}. Technological advances over
the last two decades have allowed this potential to be explored fully,
and well over 150 field studies have sought to answer this question. A
meta-analysis in 2018\textsuperscript{10} explored the pooled effect of
these studies using a meta-analytic approach to estimate species
richness of birds. Based on the combined evidence from the included
studies, they concluded that when human observers (using point counts)
and sound recorders sample areas of equal size then there is no
difference between estimates of bird species richness. When properly
conducted (see specific advice in\textsuperscript{10}), it can be
inferred that sound recorders can be used to monitor aspects of
biodiversity as efficiently as human observers. Twenty-eight primary
studies published between 2000 and 2017 were included in the
meta-analysis. Taking the role as a research funder or researcher at the
question setting stage, we can utilise cumulative meta-analysis to
determine if we need another study quantifying the difference between
acoustic recorders and human observers for bird survey point counts. We
adapted the analysis of\textsuperscript{10} to demonstrate the use of
cumulative meta-analysis (see supplementary materials). The effect size
(i.e the magnitude of the difference between intervention and control)
of studies investigating the difference between autonomous acoustic
recorders and human observers in terms of bird species richness
estimates was consistently close to 0.07 since 2015 (Figure 1). This
means that there was no clear difference between acoustic recorders and
human observers on bird point counts. It would therefore be wasteful to
fund yet another study that addressed this specific question. To reduce
research waste we need to be able to first identify it. One option is to
use cumulative meta-analysis. The approach demonstrated here is well
known and tested in the medical literature and should not be challenging
to integrate into conservation and applied ecology workflows. Cumulative
meta-analysis has already been used in our field to assess time-lag
bias\textsuperscript{11} but is not commonly used in the way we have
shown here.

\hypertarget{caveats}{%
\subsection{Caveats}\label{caveats}}

There are several important caveats that need to be addressed. The
heterogeneity in reporting and the drive for novelty in publications
means that meta-analysis is currently challenging in applied ecology.
There might not be sufficient good quality research to quantify the
cumulative effect of even some apparently well studied phenomena.
Researchers are best placed to add to the evidence base by ensuring that
they use of comparable measures of outcomes rather than novel ones.\\
In addition, publication bias, where the direction of statistical
significance of the outcome influences the decision to publish the
result, might bias the evidence base available. This is a major caveat
for all evidence synthesis approaches, but one which can be identified.
With cumulative meta-analysis one can identify publication bias11 by
accumulating the effect sizes in order of journal impact factor for
example. Although this method makes it possible to detect publication
bias it will not not solve the underlying problem, and researchers
should endeavour to reduce publication bias by following open science
(Table 1). There may of course be a time lag between identifying that we
have sufficient information, and not conducting further research on a
topic, due to timescales of publications of original research as well as
associated meta-analyses. To address these and other issues around
evidence synthesis we have been exploring a systems modelling approach
to combine empirical evidence from systematic reviews and meta-analysis
with expert opinion which allows key areas of uncertainty in a topic to
be identified and prioritised for research focus
(e.g.~\textsuperscript{12}).

\hypertarget{outlook}{%
\subsection{Outlook}\label{outlook}}

Research waste can be reduced and it is the responsibility of funders as
well as individual researchers to do so. We agree with the statement
targeted at medicine 25 years ago that ``We need\ldots{}better research,
and research done for the right reasons''\textsuperscript{13}. Without a
change in focus ecology and conservation funding will continue to be
wasted which will be detrimental to our efforts to provide solutions to
global societal challenges.

\hypertarget{supplementary-information}{%
\section{Supplementary Information}\label{supplementary-information}}

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

We extracted the data and R code from the supplementary information
in\textsuperscript{10} to recreate their analysis. As such we are
dependant on the accurate extraction of data from the primary studies by
the original authors. We intended our re-analysis of their data to be an
example of the cumulative meta-analysis approach rather than to make
explicit recommendations about the use of acoustic recorders for avian
surveys. Building on their random effects meta-analysis we ran a
cumulative meta-analysis using the ``cumul'' function in the
``metafor''\textsuperscript{14} package in R. The cumulative
meta-analysis was ordered by publication year and plotted using the
ggplot2 package in R\textsuperscript{15} . Where a single study provided
more than one estimate of effect the order in which the estimates were
accumulated was the same as the order presented by\textsuperscript{10}
and treated as subsequent trials. Changing the order of that these
particular studies were accumulated made no difference to the stability
of the estimates over time (see Figure 3). The original code, the
original data, our additional code for running the analysis and plotting
can be found at \url{https://github.com/DrMattG/Research_waste}.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-glasziou2018research}{}%
1. Glasziou, P. \& Chalmers, I. Research waste is still a scandal---an
essay by paul glasziou and iain chalmers. \emph{Bmj} \textbf{363,} k4645
(2018).

\leavevmode\hypertarget{ref-chalmers2009avoidable}{}%
2. Chalmers, I. \& Glasziou, P. Avoidable waste in the production and
reporting of research evidence. \emph{The Lancet} \textbf{374,} 86--89
(2009).

\leavevmode\hypertarget{ref-fraser2018questionable}{}%
3. Fraser, H., Parker, T., Nakagawa, S., Barnett, A. \& Fidler, F.
Questionable research practices in ecology and evolution. \emph{PloS
one} \textbf{13,} e0200303 (2018).

\leavevmode\hypertarget{ref-nilsen_bowler_linnell_2019}{}%
4. Nilsen, E. B., Bowler, D. \& Linnell, J. D. C. Exploratory and
confirmatory conservation research in the open science era. (2019).
doi:\href{https://doi.org/10.32942/osf.io/75a6f}{10.32942/osf.io/75a6f}

\leavevmode\hypertarget{ref-gurevitch2018meta}{}%
5. Gurevitch, J., Koricheva, J., Nakagawa, S. \& Stewart, G.
Meta-analysis and the science of research synthesis. \emph{Nature}
\textbf{555,} 175 (2018).

\leavevmode\hypertarget{ref-powers2019open}{}%
6. Powers, S. M. \& Hampton, S. E. Open science, reproducibility, and
transparency in ecology. \emph{Ecological applications} \textbf{29,}
e01822 (2019).

\leavevmode\hypertarget{ref-gold2013prioritizing}{}%
7. Gold, R. \emph{et al.} Prioritizing research needs based on a
systematic evidence review: A pilot process for engaging stakeholders.
\emph{Health Expectations} \textbf{16,} 338--350 (2013).

\leavevmode\hypertarget{ref-saran2018evidence}{}%
8. Saran, A. \& White, H. Evidence and gap maps: A comparison of
different approaches. \emph{Campbell Systematic Reviews} \textbf{14,}
1--38 (2018).

\leavevmode\hypertarget{ref-lau1992cumulative}{}%
9. Lau, J. \emph{et al.} Cumulative meta-analysis of therapeutic trials
for myocardial infarction. \emph{New England Journal of Medicine}
\textbf{327,} 248--254 (1992).

\leavevmode\hypertarget{ref-darras2018comparing}{}%
10. Darras, K. \emph{et al.} Comparing the sampling performance of sound
recorders versus point counts in bird surveys: A meta-analysis.
\emph{Journal of applied ecology} \textbf{55,} 2575--2586 (2018).

\leavevmode\hypertarget{ref-leimu2004cumulative}{}%
11. Leimu, R. \& Koricheva, J. Cumulative meta--analysis: A new tool for
detection of temporal trends and publication bias in ecology.
\emph{Proceedings of the Royal Society of London. Series B: Biological
Sciences} \textbf{271,} 1961--1966 (2004).

\leavevmode\hypertarget{ref-carrick2018planting}{}%
12. Carrick, J. \emph{et al.} Is planting trees the solution to reducing
flood risks? \emph{Journal of Flood Risk Management} (2018).

\leavevmode\hypertarget{ref-altman1994scandal}{}%
13. Altman, D. G. The scandal of poor medical research. (1994).

\leavevmode\hypertarget{ref-metafor}{}%
14. Viechtbauer, W. Conducting meta-analyses in R with the metafor
package. \emph{Journal of Statistical Software} \textbf{36,} 1--48
(2010).

\leavevmode\hypertarget{ref-ggplot2}{}%
15. Wickham, H. \emph{Ggplot2: Elegant graphics for data analysis}.
(Springer-Verlag New York, 2016).

\newpage

\begin{figure}
\centering
\includegraphics{MS_files/figure-latex/Add Trial-1.pdf}
\caption{Cumulative forest plot of the meta-analysis of Darras et
al.~(2018) on the difference between human observers and acoustic
recorders in terms of species richness.\label{forest}}
\end{figure}

\begin{figure}
\centering
\includegraphics{MS_files/figure-latex/flowc-1.pdf}
\caption{The production of research flows through five stages (blue
lines) all of which can lead to research waste\textsuperscript{2} (red
dashed lines). Ecology and conservation have begun to reduce waste by
focusing on methodological improvements and open science. Evidence
synthesis (including reporting to decision makers) can contribute to the
reduction in research waste by influencing question setting and
appropriate methods and design (black dashed lines).\label{flow}}
\end{figure}

\begin{figure}
\centering
\includegraphics{MS_files/figure-latex/Randomise order of the estimates in the same study-1.pdf}
\caption{Cumulative forest plot of the meta-analysis of Darras et
al.~(2018) on the difference between human observers and acoustic
recorders in terms of species richness. Estimates from the same study
inputted in random order\label{forestX}}
\end{figure}

\begin{verbatim}
## Warning: package 'huxtable' was built under R version 3.6.1
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'huxtable'
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:dplyr':
## 
##     add_rownames
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:purrr':
## 
##     every
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:scales':
## 
##     number_format
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:ggplot2':
## 
##     theme_grey
\end{verbatim}

  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
  \providecommand{\huxtpad}[1]{\rule{0pt}{\baselineskip+#1}}
  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

\begin{table}[h]
\centering
\begin{threeparttable}
\begin{tabularx}{0.35\textwidth}{p{0.175\textwidth} p{0.175\textwidth}}

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedright \textbf{Employee}\huxbpad{4pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedleft \textbf{Salary}\huxbpad{4pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{0, 0, 0}{2}}->{\huxb{0, 0, 0}{2}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedright John Smith\huxbpad{4pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedleft 50000.00\huxbpad{4pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedright Jane Doe\huxbpad{4pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedleft 50000.00\huxbpad{4pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedright David Hugh-Jones\huxbpad{4pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedleft 40000.00\huxbpad{4pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}
\end{tabularx}\end{threeparttable}

\end{table}

\end{document}

### 2019-12-02 Table1.Rmd
2.12.2019

---
title: 
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Table, echo=FALSE, message=FALSE, warning=FALSE,results="asis"}
library(tidyverse)
library(kableExtra)
df<-data.frame("Research Stage"=c("Questions relevant to stakeholders","Questions relevant to stakeholders","Appropriate design and methods","Appropriate design and methods",
                                  "Appropriate design and methods","Appropriate design and methods","Unbiased reporting","Unbiased reporting","Unbiased reporting","Unbiased reporting",
                                  "Unbiased reporting","Accessible full publication","Evidence synthesis"),
               "Examples of potential for research waste"=c("Irrelevant questions asked","Previous knowledge not properly taken into account","Study poorly designed, under-powered (or over-powered. etc.)",
                                                            "Using inappropriate statistical tools (including overfitting etc.)","Questionable research practices3 lead to poor quality research","Questionable research practices3 lead to poor quality research",
                                                            "Lack of open data","Hypothesising after the results are known","p-hacking","File-drawer syndrome (only some studies are published)","Incomplete reporting, making evidence synthesis difficult or impossible",
                                                            "Publications not available to practitioners and decision makers","Research not designed or presented in the context of the existing knowledge"),
               "Where ecology and conservation can reduce waste"=c("Co-development of research questions with stakeholders and using appropriate methodology such as Delphi exercises to avoid issues such as group think or not including the right group of experts or stakeholders",
               "Make use of evidence synthesis methods (e.g. cumulative meta-analysis, systematic mapping, systematic reviews, meta-analysis) to identify questions that are not satisfactorily answered",
               "Use simulations or power-analysis prior to undertaking data collection. Predefine effect size of interest with stakeholders (i.e. do not rely on rules of thumb for “statistical significance”)",
               "Better training of early-career researchers in methods. Open code and data to ensure reproducibility of methods",
               "Open science (open methods and data, reproducible methods, sharing code, etc.)",
               "Better training of early-career researchers in methods of open science and evidence synthesis",
               "Open science (open methods and data, reproducible methods, sharing code, etc.)",
               "Pre-registration of hypotheses", 
               "Open science (open methods and data, reproducible methods, sharing code, etc.)",
               "Pre-registration of hypotheses and methods. Open publishing (including preprints)",
               "Increasing knowledge of researchers and peer reviewers on what is essential to report, and changing journal guidelines where necessary to ensure all relevant information is reported",
               "Open access publishing, including making resources available to researchers to be able to publish open access","Using systematic reviews, systematic maps, meta-analysis, etc. to shape research priorities. Where good quality evidence is available these should be synthesised providing evidence to relevant stakeholders. Research gaps should be the focus of primary studies."
               ))

k<-kable(df,col.names = c("Research stage",
                           "Examples of potential for research waste",
                           "Where ecology and conservation can reduce waste"),booktabs = T, caption = "Table 1. The research process stages (adapted from Chalmers & Glasziou, 2009), examples of potential research waste and how ecology and conservation can limit these.",border_left = T, border_right = T) %>% column_spec(1, bold=T) %>% collapse_rows(columns = 1:3, latex_hline = "major", valign = "middle") %>% kable_styling()
#save_kable(k, "C:/Users/matthew.grainger/OneDrive - NINA/Papers_in_prep/Cumulative_meta_analysis/Research_waste/Mansuscript/Table1.pdf")

```

### 2019-12-02 Table1.tex
2.12.2019

--- 2019-12-02 Manuscript/MS.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Evidence accumulation for tackling research waste},
+            pdftitle={Table 1. The research process stages (adapted from Chalmers \& Glasziou, 2009), examples of potential research waste and how ecology and conservation can limit these.},
             pdfborder={0 0 0},
@@
 
-  \title{Evidence accumulation for tackling research waste}
+  \title{Table 1. The research process stages (adapted from Chalmers \& Glasziou,
+2009), examples of potential research waste and how ecology and
+conservation can limit these.}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
 \usepackage{array}
-\usepackage{caption}
-\usepackage{graphicx}
-\usepackage{siunitx}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
 \usepackage{colortbl}
-\usepackage{multirow}
-\usepackage{hhline}
-\usepackage{calc}
-\usepackage{tabularx}
+\usepackage{pdflscape}
+\usepackage{tabu}
 \usepackage{threeparttable}
-\usepackage{wrapfig}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-Matthew J. Grainger\textsuperscript{1}*, Frederike C.
-Bolam\textsuperscript{2}, Gavin B. Stewart\textsuperscript{2}, Erlend B.
-Nilsen\textsuperscript{1}
+\begin{tabular}{>{\bfseries}lll}
+\toprule
+Research stage & Examples of potential for research waste & Where ecology and conservation can reduce waste\\
+\midrule
+ & Irrelevant questions asked & Co-development of research questions with stakeholders and using appropriate methodology such as Delphi exercises to avoid issues such as group think or not including the right group of experts or stakeholders\\
 
-\textsuperscript{1}Norwegian Institute for Nature Research, P.O. Box
-5685 Torgarden, 7485, Trondheim, Norway
+\multirow{-2}{*}{\raggedright\arraybackslash Questions relevant to stakeholders} & Previous knowledge not properly taken into account & Make use of evidence synthesis methods (e.g.
+cumulative meta-analysis, systematic mapping, systematic reviews, meta-analysis) to identify questions that are not satisfactorily answered\\
+\cmidrule{1-3}
+ & Study poorly designed, under-powered (or over-powered.
+etc.) & Use simulations or power-analysis prior to undertaking data collection.
+Predefine effect size of interest with stakeholders (i.e.
+do not rely on rules of thumb for “statistical significance”)\\
 
-\textsuperscript{2}Modelling Evidence and Policy Research Group, School
-of Natural and Environmental Sciences, Ridley Building 2, Claremont
-Road, Newcastle upon Tyne NE1 7RU, UK
+ & Using inappropriate statistical tools (including overfitting etc.) & Better training of early-career researchers in methods. Open code and data to ensure reproducibility of methods\\
 
-*corresponding author: matthew.grainger@nina.no
+ &  & Open science (open methods and data, reproducible methods, sharing code, etc.)\\
 
-\emph{There is an urgent need for a change in research workflows so that
-pre-existing knowledge is better utilised in designing new research. A
-formal assessment of the accumulated knowledge prior to research
-approval would reduce the waste of already limited resources caused by
-asking low priority questions.}
+\multirow{-4}{*}{\raggedright\arraybackslash Appropriate design and methods} & \multirow{-2}{*}{\raggedright\arraybackslash Questionable research practices3 lead to poor quality research} & Better training of early-career researchers in methods of open science and evidence synthesis\\
+\cmidrule{1-3}
+ & Lack of open data & Open science (open methods and data, reproducible methods, sharing code, etc.)\\
 
-``Research waste'' is a well-established concept in medical
-research\textsuperscript{1}. Research is wasted when its outcomes cannot
-be used for the benefit of society\textsuperscript{2}, for example
-because no new knowledge is gained or the knowledge gained cannot be
-applied. Waste can occur at any of the four stages of the research
-process\textsuperscript{2}; question setting; methods; accessibility;
-and reporting (Table 1). In medicine, global research waste was
-estimated in 2009 to cost US \$85 bn\textsuperscript{2}, with few signs
-of improvement in the last decade\textsuperscript{1}. There is little
-reason to believe that the situation is substantially different in
-ecology and conservation, although there are no field-wide formal
-assessments of research waste.
+ & Hypothesising after the results are known & Pre-registration of hypotheses\\
 
-Emerging topics are beginning to address some of the factors that result
-in wasted research efforts (Table 1). In particular, there is increased
-focus on methodological improvements in individual studies
-(e.g.~\textsuperscript{3},\textsuperscript{4}), and on open science
-leading to improved accessibility and
-reporting\textsuperscript{5},\textsuperscript{6}. Less formal effort is
-devoted to the question setting stage. Here we suggest that ``Evidence
-Synthesis'' should be considered an additional stage of research
-(Table1, Figure S1). Evidence synthesis methods close the research
-process into a loop, and will have additional benefits in terms of
-reducing research waste at the question setting stage.
+ & p-hacking & Open science (open methods and data, reproducible methods, sharing code, etc.)\\
 
-\hypertarget{reducing-waste-in-question-setting}{%
-\section{Reducing waste in question
-setting}\label{reducing-waste-in-question-setting}}
+ & File-drawer syndrome (only some studies are published) & Pre-registration of hypotheses and methods. Open publishing (including preprints)\\
 
-There are two related areas where research waste can be reduced by
-taking into account the existing body of evidence by applying evidence
-synthesis methods.
-
-\hypertarget{low-priority-questions}{%
-\subsection{Low priority questions}\label{low-priority-questions}}
-
-New studies may ask low priority questions - those that are irrelevant
-to stakeholders. The remedy to this is to include stakeholders in the
-research commissioning process\textsuperscript{2}. Evidence synthesis,
-or horizon scanning for novel problems, should be used to provide
-evidence to practitioners, researchers and other stakeholders so that
-they can identify research gaps that are important to them and to
-develop future questions\textsuperscript{7}.
-
-\hypertarget{the-answer-is-already-known-with-certainty}{%
-\subsection{The answer is already known with
-certainty}\label{the-answer-is-already-known-with-certainty}}
-
-If a topic has been sufficiently addressed in the existing literature we
-might already know the outcome with high certainty. Further studies that
-fail to leverage this existing knowledge are at high risk of wasting
-limited research resources. There are a variety of tools available for
-research-funders and researchers to assess the state of knowledge on the
-topic of interest. For example, systematic maps (also known as Evidence
-gap maps or Evidence maps), were designed to give an overview of the
-available evidence on a broad topic\textsuperscript{8}. They can
-highlight where there is enough available evidence for a systematic
-review or where primary research is required (i.e.~there is a lack of
-evidence). Systematic reviews can be used to synthesise knowledge about
-a narrow topic such as the evidence for the effectiveness of an
-intervention and can provide a statistical summary of the pooled effect
-size. The statistical combination of numerical data extracted from the
-evidence base during the process of a systematic review is known as
-meta-analysis. Meta-analysis is commonly used in conservation and
-ecology6 providing an understanding of the magnitude of the known effect
-of an intervention across individual studies. These results can then be
-used to identify what a new research project can add to the current
-evidence base.
-
-\hypertarget{identifying-research-waste-with-cumulative-meta-analysis}{%
-\section{Identifying research waste with cumulative
-meta-analysis}\label{identifying-research-waste-with-cumulative-meta-analysis}}
-
-In medicine, one additional tool used to quantify research waste is
-cumulative meta-analysis. A cumulative meta-analysis typically describes
-the accumulation of evidence (e.g., about the effectiveness of an
-intervention) across time, and available estimates are added to the
-analysis in chronological order\textsuperscript{9}. Using cumulative
-meta-analysis, a researcher, funding agency or decision maker can
-identify if there is sufficient evidence to be confident that a reported
-effect is true. At this stage new trials are no longer required to
-predict the outcome with satisfactory certainty and hence future
-research waste will be avoided.
-
-\hypertarget{an-applied-example}{%
-\subsection{An applied example}\label{an-applied-example}}
-
-As an example of the approach within an applied ecology situation, we
-consider to what extent autonomous acoustic recorders can replace human
-observers in wildlife sampling and monitoring when the focus is on
-estimating species richness, which now has a long history in the
-ecological literature\textsuperscript{10}. Technological advances over
-the last two decades have allowed this potential to be explored fully,
-and well over 150 field studies have sought to answer this question. A
-meta-analysis in 2018\textsuperscript{10} explored the pooled effect of
-these studies using a meta-analytic approach to estimate species
-richness of birds. Based on the combined evidence from the included
-studies, they concluded that when human observers (using point counts)
-and sound recorders sample areas of equal size then there is no
-difference between estimates of bird species richness. When properly
-conducted (see specific advice in\textsuperscript{10}), it can be
-inferred that sound recorders can be used to monitor aspects of
-biodiversity as efficiently as human observers. Twenty-eight primary
-studies published between 2000 and 2017 were included in the
-meta-analysis. Taking the role as a research funder or researcher at the
-question setting stage, we can utilise cumulative meta-analysis to
-determine if we need another study quantifying the difference between
-acoustic recorders and human observers for bird survey point counts. We
-adapted the analysis of\textsuperscript{10} to demonstrate the use of
-cumulative meta-analysis (see supplementary materials). The effect size
-(i.e the magnitude of the difference between intervention and control)
-of studies investigating the difference between autonomous acoustic
-recorders and human observers in terms of bird species richness
-estimates was consistently close to 0.07 since 2015 (Figure 1). This
-means that there was no clear difference between acoustic recorders and
-human observers on bird point counts. It would therefore be wasteful to
-fund yet another study that addressed this specific question. To reduce
-research waste we need to be able to first identify it. One option is to
-use cumulative meta-analysis. The approach demonstrated here is well
-known and tested in the medical literature and should not be challenging
-to integrate into conservation and applied ecology workflows. Cumulative
-meta-analysis has already been used in our field to assess time-lag
-bias\textsuperscript{11} but is not commonly used in the way we have
-shown here.
-
-\hypertarget{caveats}{%
-\subsection{Caveats}\label{caveats}}
-
-There are several important caveats that need to be addressed. The
-heterogeneity in reporting and the drive for novelty in publications
-means that meta-analysis is currently challenging in applied ecology.
-There might not be sufficient good quality research to quantify the
-cumulative effect of even some apparently well studied phenomena.
-Researchers are best placed to add to the evidence base by ensuring that
-they use of comparable measures of outcomes rather than novel ones.\\
-In addition, publication bias, where the direction of statistical
-significance of the outcome influences the decision to publish the
-result, might bias the evidence base available. This is a major caveat
-for all evidence synthesis approaches, but one which can be identified.
-With cumulative meta-analysis one can identify publication bias11 by
-accumulating the effect sizes in order of journal impact factor for
-example. Although this method makes it possible to detect publication
-bias it will not not solve the underlying problem, and researchers
-should endeavour to reduce publication bias by following open science
-(Table 1). There may of course be a time lag between identifying that we
-have sufficient information, and not conducting further research on a
-topic, due to timescales of publications of original research as well as
-associated meta-analyses. To address these and other issues around
-evidence synthesis we have been exploring a systems modelling approach
-to combine empirical evidence from systematic reviews and meta-analysis
-with expert opinion which allows key areas of uncertainty in a topic to
-be identified and prioritised for research focus
-(e.g.~\textsuperscript{12}).
-
-\hypertarget{outlook}{%
-\subsection{Outlook}\label{outlook}}
-
-Research waste can be reduced and it is the responsibility of funders as
-well as individual researchers to do so. We agree with the statement
-targeted at medicine 25 years ago that ``We need\ldots{}better research,
-and research done for the right reasons''\textsuperscript{13}. Without a
-change in focus ecology and conservation funding will continue to be
-wasted which will be detrimental to our efforts to provide solutions to
-global societal challenges.
-
-\hypertarget{supplementary-information}{%
-\section{Supplementary Information}\label{supplementary-information}}
-
-\hypertarget{methods}{%
-\subsection{Methods}\label{methods}}
-
-We extracted the data and R code from the supplementary information
-in\textsuperscript{10} to recreate their analysis. As such we are
-dependant on the accurate extraction of data from the primary studies by
-the original authors. We intended our re-analysis of their data to be an
-example of the cumulative meta-analysis approach rather than to make
-explicit recommendations about the use of acoustic recorders for avian
-surveys. Building on their random effects meta-analysis we ran a
-cumulative meta-analysis using the ``cumul'' function in the
-``metafor''\textsuperscript{14} package in R. The cumulative
-meta-analysis was ordered by publication year and plotted using the
-ggplot2 package in R\textsuperscript{15} . Where a single study provided
-more than one estimate of effect the order in which the estimates were
-accumulated was the same as the order presented by\textsuperscript{10}
-and treated as subsequent trials. Changing the order of that these
-particular studies were accumulated made no difference to the stability
-of the estimates over time (see Figure 3). The original code, the
-original data, our additional code for running the analysis and plotting
-can be found at \url{https://github.com/DrMattG/Research_waste}.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-glasziou2018research}{}%
-1. Glasziou, P. \& Chalmers, I. Research waste is still a scandal---an
-essay by paul glasziou and iain chalmers. \emph{Bmj} \textbf{363,} k4645
-(2018).
-
-\leavevmode\hypertarget{ref-chalmers2009avoidable}{}%
-2. Chalmers, I. \& Glasziou, P. Avoidable waste in the production and
-reporting of research evidence. \emph{The Lancet} \textbf{374,} 86--89
-(2009).
-
-\leavevmode\hypertarget{ref-fraser2018questionable}{}%
-3. Fraser, H., Parker, T., Nakagawa, S., Barnett, A. \& Fidler, F.
-Questionable research practices in ecology and evolution. \emph{PloS
-one} \textbf{13,} e0200303 (2018).
-
-\leavevmode\hypertarget{ref-nilsen_bowler_linnell_2019}{}%
-4. Nilsen, E. B., Bowler, D. \& Linnell, J. D. C. Exploratory and
-confirmatory conservation research in the open science era. (2019).
-doi:\href{https://doi.org/10.32942/osf.io/75a6f}{10.32942/osf.io/75a6f}
-
-\leavevmode\hypertarget{ref-gurevitch2018meta}{}%
-5. Gurevitch, J., Koricheva, J., Nakagawa, S. \& Stewart, G.
-Meta-analysis and the science of research synthesis. \emph{Nature}
-\textbf{555,} 175 (2018).
-
-\leavevmode\hypertarget{ref-powers2019open}{}%
-6. Powers, S. M. \& Hampton, S. E. Open science, reproducibility, and
-transparency in ecology. \emph{Ecological applications} \textbf{29,}
-e01822 (2019).
-
-\leavevmode\hypertarget{ref-gold2013prioritizing}{}%
-7. Gold, R. \emph{et al.} Prioritizing research needs based on a
-systematic evidence review: A pilot process for engaging stakeholders.
-\emph{Health Expectations} \textbf{16,} 338--350 (2013).
-
-\leavevmode\hypertarget{ref-saran2018evidence}{}%
-8. Saran, A. \& White, H. Evidence and gap maps: A comparison of
-different approaches. \emph{Campbell Systematic Reviews} \textbf{14,}
-1--38 (2018).
-
-\leavevmode\hypertarget{ref-lau1992cumulative}{}%
-9. Lau, J. \emph{et al.} Cumulative meta-analysis of therapeutic trials
-for myocardial infarction. \emph{New England Journal of Medicine}
-\textbf{327,} 248--254 (1992).
-
-\leavevmode\hypertarget{ref-darras2018comparing}{}%
-10. Darras, K. \emph{et al.} Comparing the sampling performance of sound
-recorders versus point counts in bird surveys: A meta-analysis.
-\emph{Journal of applied ecology} \textbf{55,} 2575--2586 (2018).
-
-\leavevmode\hypertarget{ref-leimu2004cumulative}{}%
-11. Leimu, R. \& Koricheva, J. Cumulative meta--analysis: A new tool for
-detection of temporal trends and publication bias in ecology.
-\emph{Proceedings of the Royal Society of London. Series B: Biological
-Sciences} \textbf{271,} 1961--1966 (2004).
-
-\leavevmode\hypertarget{ref-carrick2018planting}{}%
-12. Carrick, J. \emph{et al.} Is planting trees the solution to reducing
-flood risks? \emph{Journal of Flood Risk Management} (2018).
-
-\leavevmode\hypertarget{ref-altman1994scandal}{}%
-13. Altman, D. G. The scandal of poor medical research. (1994).
-
-\leavevmode\hypertarget{ref-metafor}{}%
-14. Viechtbauer, W. Conducting meta-analyses in R with the metafor
-package. \emph{Journal of Statistical Software} \textbf{36,} 1--48
-(2010).
-
-\leavevmode\hypertarget{ref-ggplot2}{}%
-15. Wickham, H. \emph{Ggplot2: Elegant graphics for data analysis}.
-(Springer-Verlag New York, 2016).
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{MS_files/figure-latex/Add Trial-1.pdf}
-\caption{Cumulative forest plot of the meta-analysis of Darras et
-al.~(2018) on the difference between human observers and acoustic
-recorders in terms of species richness.\label{forest}}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{MS_files/figure-latex/flowc-1.pdf}
-\caption{The production of research flows through five stages (blue
-lines) all of which can lead to research waste\textsuperscript{2} (red
-dashed lines). Ecology and conservation have begun to reduce waste by
-focusing on methodological improvements and open science. Evidence
-synthesis (including reporting to decision makers) can contribute to the
-reduction in research waste by influencing question setting and
-appropriate methods and design (black dashed lines).\label{flow}}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{MS_files/figure-latex/Randomise order of the estimates in the same study-1.pdf}
-\caption{Cumulative forest plot of the meta-analysis of Darras et
-al.~(2018) on the difference between human observers and acoustic
-recorders in terms of species richness. Estimates from the same study
-inputted in random order\label{forestX}}
-\end{figure}
-
-\begin{verbatim}
-## Warning: package 'huxtable' was built under R version 3.6.1
-\end{verbatim}
-
-\begin{verbatim}
-## 
-## Attaching package: 'huxtable'
-\end{verbatim}
-
-\begin{verbatim}
-## The following object is masked from 'package:dplyr':
-## 
-##     add_rownames
-\end{verbatim}
-
-\begin{verbatim}
-## The following object is masked from 'package:purrr':
-## 
-##     every
-\end{verbatim}
-
-\begin{verbatim}
-## The following object is masked from 'package:scales':
-## 
-##     number_format
-\end{verbatim}
-
-\begin{verbatim}
-## The following object is masked from 'package:ggplot2':
-## 
-##     theme_grey
-\end{verbatim}
-
-  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
-  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
-  \providecommand{\huxtpad}[1]{\rule{0pt}{\baselineskip+#1}}
-  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
-
-\begin{table}[h]
-\centering
-\begin{threeparttable}
-\begin{tabularx}{0.35\textwidth}{p{0.175\textwidth} p{0.175\textwidth}}
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedright \textbf{Employee}\huxbpad{4pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedleft \textbf{Salary}\huxbpad{4pt}} \tabularnewline[-0.5pt]
-
-\hhline{>{\huxb{0, 0, 0}{2}}->{\huxb{0, 0, 0}{2}}-}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedright John Smith\huxbpad{4pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedleft 50000.00\huxbpad{4pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedright Jane Doe\huxbpad{4pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedleft 50000.00\huxbpad{4pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedright David Hugh-Jones\huxbpad{4pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{4pt}\raggedleft 40000.00\huxbpad{4pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-\end{tabularx}\end{threeparttable}
-
-\end{table}
+\multirow{-5}{*}{\raggedright\arraybackslash Unbiased reporting} & Incomplete reporting, making evidence synthesis difficult or impossible & Increasing knowledge of researchers and peer reviewers on what is essential to report, and changing journal guidelines where necessary to ensure all relevant information is reported\\
+\cmidrule{1-3}
+Accessible full publication & Publications not available to practitioners and decision makers & Open access publishing, including making resources available to researchers to be able to publish open access\\
+\cmidrule{1-3}
+Evidence synthesis & Research not designed or presented in the context of the existing knowledge & Using systematic reviews, systematic maps, meta-analysis, etc.
+to shape research priorities.
+Where good quality evidence is available these should be synthesised providing evidence to relevant stakeholders.
+Research gaps should be the focus of primary studies.\\
+\bottomrule
+\end{tabular}
 

### 2019-12-02 Manuscript/MS.Rmd
Spell Rikes name correctly!

@@
 
-Matthew J. Grainger^1^*, Frederike C. Bolam^2^, Gavin B. Stewart^2^, Erlend B. Nilsen^1^
+Matthew J. Grainger^1^*, Friederike C. Bolam^2^, Gavin B. Stewart^2^, Erlend B. Nilsen^1^
 
@@
 
+Table 1 can be found at: https://github.com/DrMattG/Research_waste/blob/master/Manuscript/Table1.pdf
+
 ```{r, echo=FALSE}
+
 # df<-data.frame("Research Stage"=c("Questions relevant to stakeholders","Questions relevant to stakeholders","Appropriate design and methods","Appropriate design and methods",
@@
 # df %>% kable("latex") %>% as_image()
-library(huxtable)
-ht <- hux(
-  Employee = c('John Smith', 'Jane Doe', 'David Hugh-Jones'),
-  Salary = c(50000, 50000, 40000),
-  add_colnames = TRUE
-)
-
-bold(ht)[1,]           <- TRUE
-bottom_border(ht)[1,]  <- 2
-align(ht)[,2]          <- 'right'
-right_padding(ht)      <- 10
-left_padding(ht)       <- 10
-width(ht)              <- 0.35
-number_format(ht)      <- 2
-
-ht
 

### 2019-12-16 Manuscript/MS.Rmd
Revision1_Final_edits

@@
 
-# Supplementary Information
+*There is an urgent need for a change in research workflows so that pre-existing knowledge is better utilised in designing new research.
+A formal assessment of the accumulated knowledge prior to research approval would reduce the waste of already limited resources caused by asking low priority questions.*
 
-## Methods
-We extracted the data and R code from the supplementary information in [@darras2018comparing] to recreate their analysis.
-As such we are dependant on the accurate extraction of data from the primary studies by the original authors.
-We intended our re-analysis of their data to be an example of the cumulative meta-analysis approach rather than to make explicit recommendations about the use of acoustic recorders for avian surveys.
-Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” [@metafor] package in R.
-The cumulative meta-analysis was ordered by publication year and plotted using the ggplot2 package in R [@ggplot2] .
-Where a single study provided more than one estimate of effect the order in which the estimates were accumulated was the same as the order presented by [@darras2018comparing] and treated as subsequent trials.
-Changing the order of that these particular studies were accumulated made no difference to the stability of the estimates over time (see Figure 3).
-The original code, the original data, our additional code for running the analysis and plotting can be found at https://github.com/DrMattG/Research_waste.
+“Research waste” is a well-established concept in medical research[@glasziou2018research].
+Research is wasted when its outcomes cannot be used for the benefit of society[@chalmers2009avoidable], for the benefit of training students or the benefit of engaging stakeholders, for example because no new knowledge is gained or the knowledge gained cannot be applied.
+Waste can occur at any of the four stages of the research process[@chalmers2009avoidable]; question setting; methods; accessibility; and reporting (Table 1).
+In medicine, global research waste was estimated in 2009 to cost US$85bn[@chalmers2009avoidable], with few signs of improvement in the last decade[@glasziou2018research].
+There is little reason to believe that the situation is substantially different in ecology and conservation, although there are no field-wide formal assessments of research waste.
+
+Emerging topics are beginning to address some of the factors that result in wasted research efforts (Table 1).
+In particular, there is increased focus on methodological improvements in individual studies (e.g.
+[@fraser2018questionable], [@nilsen_bowler_linnell_2019]),  and on open science leading to improved accessibility and reporting[@powers2019open],  [@gurevitch2018meta].
+Less formal effort is devoted to the question setting stage.
+Here we suggest that “Evidence Synthesis” should be considered an additional stage of research (Table1, Figure S1).
+Evidence synthesis methods close the research process into a loop, and will have additional benefits in terms of reducing research waste at the question setting stage.
+
+# Reducing waste in question setting
+There are two related areas where research waste can be reduced by taking into account the existing body of evidence by applying evidence synthesis methods.
+
+## Low priority questions
+
+New studies may ask low priority questions - those that are irrelevant to stakeholders.
+The remedy to this is to include stakeholders in the research commissioning process[@chalmers2009avoidable].
+Evidence synthesis, or horizon scanning for novel problems, should be used to provide evidence to practitioners, researchers and other stakeholders so that they can identify research gaps that are important to them and to develop future questions[@gold2013prioritizing].
+
+## The answer is already known with certainty
+
+If a topic has been sufficiently addressed in the existing literature we might already know the outcome with high certainty.
+Further studies that fail to leverage this existing knowledge are at high risk of wasting limited research resources.
+There are a variety of tools available for research-funders and researchers to assess the state of knowledge on the topic of interest.
+For example, systematic maps (also known as Evidence gap maps or Evidence maps), were designed to give an overview of the available evidence on a broad topic[@saran2018evidence].
+They can highlight where there is enough available evidence for a systematic review or where primary research is required (i.e.
+there is a lack of evidence).
+However, users of systematic maps should be aware that the number of papers available on a topic does not equate to the strength of evidence which should be formally examined before making conclusions about if sufficient evidence is available(@stewart2015use].
+Systematic reviews can be used to synthesise knowledge about a narrow topic such as the evidence for the effectiveness of an intervention and can provide a statistical summary of the pooled effect size.
+The statistical combination of numerical data extracted from the evidence base during the process of a systematic review is known as meta-analysis.
+Meta-analysis is commonly used in conservation and ecology[@gurevitch2018meta] providing an understanding of the magnitude of the known effect of an intervention across individual studies.
+These results can then be used to identify what a new research project can add to the current evidence base.
+
+## Identifying research waste with cumulative meta-analysis
+
+In medicine, one additional tool used to quantify research waste is cumulative meta-analysis.
+A cumulative meta-analysis typically describes the accumulation of evidence (e.g., about the effectiveness of an intervention) across time, and available estimates are added to the analysis in chronological order[@lau1992cumulative].
+Using cumulative meta-analysis, a researcher, funding agency or decision maker can identify if there is sufficient evidence to be confident that a reported effect is true.
+At this stage new trials are no longer required to predict the outcome with satisfactory certainty and hence future research waste will be avoided.
+Cumulative meta-analyses demonstrate how new research is frequently undertaken generating research waste, even when effect sizes are temporally stable and precise[@lau1992cumulative].
+Researchers in domains relying on heterogeneous observational studies (such as ecologists) should beware of temporal instability of effects[@koricheva2019temporal] which should be considered as part of the strength of the existing evidence-base.
+
+## An applied example
+
+As an example, we consider to what extent autonomous acoustic recorders can replace human observers in wildlife sampling and monitoring when the focus is on estimating species richness, which now has a long history in the ecological literature12.
+Technological advances over the last two decades have allowed this potential to be explored fully, and well over 150 field studies have sought to answer this question.
+A meta-analysis in 2018[@darras2018comparing] explored the pooled effect of these studies using a meta-analytic approach to estimate species richness of birds.
+Based on the combined evidence from the included studies, they concluded that when human observers (using point counts) and sound recorders sample areas of equal size then there is no difference between estimates of bird species richness.
+When properly conducted (see specific advice in [@darras2018comparing]), it can be inferred that sound recorders can be used to monitor aspects of biodiversity as efficiently as human observers.
+Twenty-eight primary studies published between 2000 and 2017 were included in the meta-analysis.
+Taking the role as a research funder or researcher at the question setting stage, we can utilise cumulative meta-analysis to determine if we need another study quantifying the difference between acoustic recorders and human observers for bird survey point counts.
+We adapted the analysis of [@darras2018comparing] to demonstrate the use of cumulative meta-analysis (see supplementary materials).
+The effect size (i.e the magnitude of the difference between intervention and control) of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0.07 since 2015 (Figure 1).
+This means that there was no clear difference between acoustic recorders and human observers on bird point counts.
+It would be wasteful to fund yet another study that addressed this specific question.
+To reduce research waste we need to be able to first identify it.
+One option is to use cumulative meta-analysis.
+The approach demonstrated here is well known and tested in the medical literature and should not be challenging to integrate into conservation and applied ecology workflows.
+Cumulative meta-analysis has already been used in our field to assess time-lag bias[@leimu2004cumulative] but is not commonly used in the way we have shown here.
+
+## Caveats
+
+There are several important caveats that need to be addressed.
+The heterogeneity in reporting and the drive for novelty in publications means that meta-analysis is currently challenging in applied ecology.
+There might not be sufficient good quality research to quantify the cumulative effect of even some apparently well studied phenomena.
+Researchers are best placed to add to the evidence base by ensuring that they use of comparable measures of outcomes rather than novel ones.
+In addition, publication bias, where the direction of statistical significance of the outcome influences the decision to publish the result, might bias the evidence base available.
+This is a major caveat for all evidence synthesis approaches, but one which can be identified.
+With cumulative meta-analysis one can explore publication bias [@leimu2004cumulative] by accumulating the effect sizes in order of journal impact factor for example.
+Although this method makes it possible to detect publication bias it will not solve the underlying problem, and researchers should endeavour to reduce publication bias by following open science (Table 1).
+To address this and the problem of synthesising diverse information sources in non-linear systems with multiple complexities, methodologists propose use of systems models to combine empirical evidence from systematic reviews and meta-analysis with expert opinion which allows key areas of uncertainty in a topic to be identified and prioritised for research focus (e.g.
+[@carrick2018planting]).
+Formal value of information analysis can then be undertaken if a decision-theoretic framework exists.
+
+## Outlook
+Research waste can be reduced and it is the responsibility of funders as well as individual researchers to do so.
+Researchers and funders could search for existing research syntheses in the literature and on synthesis platforms (e.g.
+https://www.conservationevidence.com/).
+We agree with the statement targeted at medicine 25 years ago that “We need...better research, and research done for the right reasons” [@altman1994scandal].
+Without a change in focus ecology and conservation funding will continue to be wasted which will be detrimental to our efforts to provide solutions to global societal challenges.
 
@@
 
-```{r flowc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "The production of research flows through five stages (blue lines) all of which can lead to research waste [@@chalmers2009avoidable] (red dashed lines).
+```{r flowc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "The production of research flows through five stages (blue lines) all of which can lead to research waste[@@chalmers2009avoidable] (red dashed lines).
 Ecology and conservation have begun to reduce waste by focusing on methodological improvements and open science.
-Evidence synthesis (including reporting to decision makers) can contribute to the reduction in research waste by influencing question setting and appropriate methods and design (black dashed lines).\\label{flow}"}
+Evidence synthesis (including reporting to decision makers) can contribute to the reduction in research waste by influencing question setting and appropriate methods and design (black dashed lines).
+Poor evidence synthesis can also lead to research waste.\\label{flow}"}
+
 library(DiagrammeR)
@@
 
-```{r Randomise order of the estimates in the same study,echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
-(2018) on the difference between human observers and acoustic recorders in terms of species richness.
-Estimates from the same study inputted in random order\\label{forestX}"}
-set.seed(8955)
-metax<-meta1
-metax$order<-order(metax$year, runif(length(metax$year))) 
-res2 <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=metax, slab=paste(metax$first_author, metax$year, sep=","))
-cres2<-cumul(res2, order=order(metax$order))
-par(mfrow=c(1,2))
-forest(cres2)
-forest(cres)
-
-```
-
 Table 1 can be found at: https://github.com/DrMattG/Research_waste/blob/master/Manuscript/Table1.pdf

### 2019-12-16 Manuscript/SuppMat.Rmd
Revision1_Final_edits

---
title: "Evidence accumulation for tackling research waste (Supplement)"
csl: nature.csl
output:
  pdf_document:
    fig_caption: yes
  word_document: 
    fig_caption: yes
  html_document: 
    fig_caption: yes
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Matthew J. Grainger^1^*, Friederike C. Bolam^2^, Gavin B. Stewart^2^, Erlend B. Nilsen^1^

^1^Norwegian Institute for Nature Research, P.O. Box 5685 Torgarden, 7485, Trondheim, Norway

^2^Modelling Evidence and Policy Research Group, School of Natural and Environmental Sciences, Ridley Building 2, Claremont Road, Newcastle upon Tyne NE1 7RU, UK

*corresponding author: matthew.grainger\@nina.no

## Supplementary Information

# Methods
We extracted the data and R code from the supplementary information in [@darras2018comparing] to recreate their analysis. As such we are dependent on the accurate extraction of data from the primary studies by the original authors. We intended our re-analysis of their data to be an example of the cumulative meta-analysis approach rather than to make explicit recommendations about the use of acoustic recorders for avian surveys. Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” [@metafor] package in R. The cumulative meta-analysis was ordered by publication year and plotted using the ggplot2 package in R[@ggplot2]. Where a single study provided more than one estimate of effect the order in which the estimates were accumulated was the same as the order presented by [@darras2018comparing] and treated as subsequent trials. Changing the order of that these particular studies were accumulated made no difference to the stability of the estimates over time (see the code for an assessment). The original code, the original data, our additional code for running the analysis and plotting can be found at https://github.com/DrMattG/Research_waste.

# References {-}

<div id="refs"></div>

\newpage

### 2020-01-27 Manuscript/MS.Rmd
2nd round review update 2

--- 2019-12-02 Manuscript/MS.Rmd
@@
 
-*There is an urgent need for a change in research workflows so that pre-existing knowledge is better utilised in designing new research.
+*There is an immediate need for a change in research workflows so that pre-existing knowledge is better utilised in designing new research.
 A formal assessment of the accumulated knowledge prior to research approval would reduce the waste of already limited resources caused by asking low priority questions.*
 
-“Research waste” is a well-established concept in medical research[@glasziou2018research].
-Research is wasted when its outcomes cannot be used for the benefit of society [@chalmers2009avoidable], for example because no new knowledge is gained or the knowledge gained cannot be applied.
-Waste can occur at any of the four stages of the research process [@chalmers2009avoidable]; question setting; methods; accessibility; and reporting (Table 1).
-In medicine, global research waste was estimated in 2009 to cost US $85 bn [@chalmers2009avoidable], with few signs of improvement in the last decade [@glasziou2018research].
+“Research waste” is defined as research outcomes with no societal benefits [@glasziou2018research].
+This occurs when the knowledge generated is either not innovative and novel, when it is not applicable to real-world situations, or when the research effort itself does not have societal impact (e.g.
+training students or engaging stakeholders).
+Waste can occur at any stage of the research process [@chalmers2009avoidable]; question setting; methods; accessibility; and reporting (Table 1).
+In medicine, research waste is a well-established concept, and in 2009 research waste was estimated to cost US$85bn [@chalmers2009avoidable], with few signs of improvement in the last decade [@glasziou2018research].
 There is little reason to believe that the situation is substantially different in ecology and conservation, although there are no field-wide formal assessments of research waste.
@@
 In particular, there is increased focus on methodological improvements in individual studies (e.g.
-[@fraser2018questionable], [@nilsen_bowler_linnell_2019]),  and on open science leading to improved accessibility and reporting [@gurevitch2018meta], [@powers2019open].
+[@fraser2018questionable], [@nilsen_bowler_linnell_2019]),  and on open science leading to improved accessibility and reporting [@gurevitch2018meta].
 Less formal effort is devoted to the question setting stage.
-Here we suggest that “Evidence Synthesis” should be considered an additional stage of research (Table1, Figure S1).
-Evidence synthesis methods close the research process into a loop, and will have additional benefits in terms of reducing research waste at the question setting stage.
+Here we suggest that “Evidence Synthesis” should be considered an additional stage of research (Table1, Supplementary Figure 1)  which closes the  research process into a loop,  leading to additional benefits in terms of reducing research waste at the question setting stage.
 
 # Reducing waste in question setting
-
-There are two related areas where research waste can be reduced by taking into account the existing body of evidence by applying evidence synthesis methods.
+There are two related areas where research waste can be reduced by taking into account the existing body of evidence through evidence synthesis methods.
 
@@
 New studies may ask low priority questions - those that are irrelevant to stakeholders.
-The remedy to this is to include stakeholders in the research commissioning process[@chalmers2009avoidable].
+The remedy to this is to include stakeholders in the research commissioning process [@chalmers2009avoidable].
 Evidence synthesis, or horizon scanning for novel problems, should be used to provide evidence to practitioners, researchers and other stakeholders so that they can identify research gaps that are important to them and to develop future questions [@gold2013prioritizing].
@@
 If a topic has been sufficiently addressed in the existing literature we might already know the outcome with high certainty.
-Further studies that fail to leverage this existing knowledge are at high risk of wasting limited research resources.
+Further studies that fail to leverage this existing knowledge are at risk of wasting research resources.
 There are a variety of tools available for research-funders and researchers to assess the state of knowledge on the topic of interest.
 For example, systematic maps (also known as Evidence gap maps or Evidence maps), were designed to give an overview of the available evidence on a broad topic [@saran2018evidence].
-They can highlight where there is enough available evidence for a systematic review or where primary research is required (i.e.
-there is a lack of evidence).
+They can highlight where there is enough available evidence for a systematic review or where primary research is required due to a lack of evidence.
+However, users of systematic maps should be aware that the number of papers available on a topic does not equate to the strength of evidence which should be formally examined before making conclusions about the adequacy of evidence [@stewart2015use].
 Systematic reviews can be used to synthesise knowledge about a narrow topic such as the evidence for the effectiveness of an intervention and can provide a statistical summary of the pooled effect size.
 The statistical combination of numerical data extracted from the evidence base during the process of a systematic review is known as meta-analysis.
-Meta-analysis is commonly used in conservation and ecology6 providing an understanding of the magnitude of the known effect of an intervention across individual studies.
-These results can then be used to identify what a new research project can add to the current evidence base.
+Meta-analysis is increasingly used in conservation and ecology [@gurevitch2018meta] providing an understanding of the magnitude of the known effect of an intervention across individual studies.
+These results can be used to identify what new research can add to the current evidence base.
+Although evidence synthesis can be time consuming (opensource tools for predicting the time investment e.g.
+https://github.com/mjwestgate/PredicTER are available) the investment in time will facilitate less wasteful research.
 
-# Identifying research waste with cumulative meta-analysis
+## Identifying research waste with cumulative meta-analysis
+
 In medicine, one additional tool used to quantify research waste is cumulative meta-analysis.
 A cumulative meta-analysis typically describes the accumulation of evidence (e.g., about the effectiveness of an intervention) across time, and available estimates are added to the analysis in chronological order [@lau1992cumulative].
-Using cumulative meta-analysis, a researcher, funding agency or decision maker can identify if there is sufficient evidence to be confident that a reported effect is true.
-At this stage new trials are no longer required to predict the outcome with satisfactory certainty and hence future research waste will be avoided.
-
-## An applied example
-As an example of the approach within an applied ecology situation, we consider to what extent autonomous acoustic recorders can replace human observers in wildlife sampling and monitoring when the focus is on estimating species richness, which now has a long history in the ecological literature [@darras2018comparing].
-Technological advances over the last two decades have allowed this potential to be explored fully, and well over 150 field studies have sought to answer this question.
-A meta-analysis in 2018[@darras2018comparing] explored the pooled effect of these studies using a meta-analytic approach to estimate species richness of birds.
-Based on the combined evidence from the included studies, they concluded that when human observers (using point counts) and sound recorders sample areas of equal size then there is no difference between estimates of bird species richness.
-When properly conducted (see specific advice in [@darras2018comparing]), it can be inferred that sound recorders can be used to monitor aspects of biodiversity as efficiently as human observers.
-Twenty-eight primary studies published between 2000 and 2017 were included in the meta-analysis.
-Taking the role as a research funder or researcher at the question setting stage, we can utilise cumulative meta-analysis to determine if we need another study quantifying the difference between acoustic recorders and human observers for bird survey point counts.
-We adapted the analysis of [@darras2018comparing] to demonstrate the use of cumulative meta-analysis (see supplementary materials).
-The effect size (i.e the magnitude of the difference between intervention and control) of studies investigating the difference between autonomous acoustic recorders and human observers in terms of bird species richness estimates was consistently close to 0.07 since 2015 (Figure 1).
-This means that there was no clear difference between acoustic recorders and human observers on bird point counts.
-It would therefore be wasteful to fund yet another study that addressed this specific question.
-To reduce research waste we need to be able to first identify it.
-One option is to use cumulative meta-analysis.
-The approach demonstrated here is well known and tested in the medical literature and should not be challenging to integrate into conservation and applied ecology workflows.
-Cumulative meta-analysis has already been used in our field to assess time-lag bias [@leimu2004cumulative] but is not commonly used in the way we have shown here.
+Using cumulative meta-analysis, one can identify if there is sufficient evidence to be confident that a reported effect is true.
+Cumulative meta-analyses demonstrate how new research is frequently undertaken generating research waste, even when effect sizes are temporally stable and precise [@lau1992cumulative].
+Researchers in domains relying on heterogeneous observational studies (such as ecologists) should beware of temporal instability of effects [@koricheva2019temporal] which should be considered as part of the strength of the existing evidence-base.
+We demonstrate how the approach can be integrated into conservation and applied ecology workflows,  see Box 1.
+Cumulative meta-analysis has already been used in our field to assess time-lag bias [@leimu2004cumulative] but is not commonly used in the way we show here.
 
 ## Caveats
+
 There are several important caveats that need to be addressed.
@@
 There might not be sufficient good quality research to quantify the cumulative effect of even some apparently well studied phenomena.
-Researchers are best placed to add to the evidence base by ensuring that they use of comparable measures of outcomes rather than novel ones.
+Power analysis can be used to determine if there is enough primary literature to determine stability of a cumulative effect.
+Researchers are best placed to add to the evidence base by ensuring that they use comparable measures of outcomes rather than novel ones.
 In addition, publication bias, where the direction of statistical significance of the outcome influences the decision to publish the result, might bias the evidence base available.
 This is a major caveat for all evidence synthesis approaches, but one which can be identified.
-With cumulative meta-analysis one can identify publication bias11 by accumulating the effect sizes in order of journal impact factor for example.
-Although this method makes it possible to detect publication bias it will not not solve the underlying problem, and researchers should endeavour to reduce publication bias by following open science (Table 1).
-There may of course be a time lag between identifying that we have sufficient information, and not conducting further research on a topic, due to timescales of publications of original research as well as associated meta-analyses.
-To address these and other issues around evidence synthesis we have been exploring a systems modelling approach to combine empirical evidence from systematic reviews and meta-analysis with expert opinion which allows key areas of uncertainty in a topic to be identified and prioritised for research focus (e.g.
+Funnel plots can be used to identify the potential for non-publication of results (i.e.
+those of small effect size).
+With cumulative meta-analysis one can explore publication bias [@leimu2004cumulative] by accumulating the effect sizes in order of journal impact factor for example.
+Although this method makes it possible to detect publication bias it will not solve the underlying problem, and researchers should endeavour to reduce publication bias by following open science principles (Table 1).
+Without deliberate research programs of sequential research focused on specific questions in ecology, as proposed in [@nichols2019accumulating], we need to find ways of  synthesising diverse information sources.
+Methodologists propose use of systems models to combine empirical evidence from systematic reviews and meta-analysis with expert opinion which allows key areas of uncertainty in a topic to be identified and prioritised for research focus (e.g.
 [@carrick2018planting]).
+Formal value of information analysis can then be undertaken if a decision-theoretic framework exists.
 
 ## Outlook
-Research waste can be reduced and it is the responsibility of funders as well as individual researchers to do so.
+
+Research waste can be reduced and it is the responsibility of research funders, research institutions as well as individual researchers to do so.
+Researchers and funders could search for existing research syntheses in the literature and on synthesis platforms (e.g.
+https://www.conservationevidence.com/).
 We agree with the statement targeted at medicine 25 years ago that “We need...better research, and research done for the right reasons” [@altman1994scandal].
@@
 
-# Supplementary Information
+### Box 1 Using cumulative meta-analysis to make research decisions
 
-## Methods
-We extracted the data and R code from the supplementary information in [@darras2018comparing] to recreate their analysis.
-As such we are dependant on the accurate extraction of data from the primary studies by the original authors.
-We intended our re-analysis of their data to be an example of the cumulative meta-analysis approach rather than to make explicit recommendations about the use of acoustic recorders for avian surveys.
-Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” [@metafor] package in R.
-The cumulative meta-analysis was ordered by publication year and plotted using the ggplot2 package in R [@ggplot2] .
-Where a single study provided more than one estimate of effect the order in which the estimates were accumulated was the same as the order presented by [@darras2018comparing] and treated as subsequent trials.
-Changing the order of that these particular studies were accumulated made no difference to the stability of the estimates over time (see Figure 3).
-The original code, the original data, our additional code for running the analysis and plotting can be found at https://github.com/DrMattG/Research_waste.
+Imagine you, as a researcher or research funder, want to assess the potential for acoustic recorders to replace human observers for estimating bird abundance.
+Do we need another research study to determine this?
+Here we outline an example decision process which would serve to reduce research waste.
+
+* **Is there previous knowledge available on this topic?** Technological advances over the last two decades have allowed this potential to be explored fully, and well over 150 field studies have sought to answer this question.
+A meta-analysis in 201815 explored the pooled effect of these studies using a meta-analytic approach to estimate species richness of birds.
+Twenty-eight primary studies published between 2000 and 2017 were included in the meta-analysis.
+Based on the combined evidence from the included studies, they concluded that when human observers (using point counts) and sound recorders sample areas of equal size then there is no difference between estimates of bird species richness.
+When properly conducted (see specific advice in [@darras2018comparing]), it can be inferred that sound recorders can be used to monitor aspects of biodiversity as efficiently as human observers.
+*	**Do we need another study to quantify the effect?** We adapted the analysis of [@darras2018comparing] to demonstrate the use of cumulative meta-analysis (see Supplementary Methods).
+For the 14 studies conducted since 2014, the effect size (i.e.
+the magnitude of the difference between intervention and control) of studies was consistently close to 0.07  (a difference of around 1 species; Figure 1).
+This means that there was no clear difference between acoustic recorders and human observers on bird point counts.
+It would be wasteful to fund yet another study that addressed this specific question.
+*	**What next?** If you were interested in acoustic sampling as a means to sample bird species richness you could proceed with confidence that using acoustic sampling is as effective as human observers in the field.
+If you are interested in acoustic sampling specifically you could look for substantial anomalies or heterogeneity between studies in the meta-analysis and design a study trying to understand these differences.
 
@@
 ```{r Add Trial, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
-(2018) on the difference between human observers and acoustic recorders in terms of species richness.\\label{forest}"}
+(2018) on the difference between human observers and acoustic recorders in terms of species richness.
+The green line indicates the line of zero effect, the blue dots indicate the cumulative effect size with 95% confidence intervals.
+Studies are ordered by publication year with the earliest studies at the top and the latest studies at the bottom.\\label{forest}"}
 ### Add cumulative meta-analysis
@@
 
-```
-
-```{r flowc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "The production of research flows through five stages (blue lines) all of which can lead to research waste [@@chalmers2009avoidable] (red dashed lines).
-Ecology and conservation have begun to reduce waste by focusing on methodological improvements and open science.
-Evidence synthesis (including reporting to decision makers) can contribute to the reduction in research waste by influencing question setting and appropriate methods and design (black dashed lines).\\label{flow}"}
-library(DiagrammeR)
-
-grViz("
-digraph boxes_and_circles {
-
-  # a 'graph' statement
-  graph [overlap = true, fontsize = 10]
-
-  # several 'node' statements
-  node [shape = box,
-        fontname = Helvetica]
-      tab1 [label = '@@1',color = blue]
-      tab2 [label = '@@2',color = blue]
-      tab3 [label = '@@3',color = blue]
-      tab4 [label = '@@4',color = blue]
-      tab6 [label = '@@6',color = blue]
-      tab8 [label = '@@8',color = red]
-      tab9 [label = '@@9',color = red]
-      tab10[label = '@@10',color = red]
-      tab11[label = '@@11',color = red]
-
-  node [shape = circle,
-        fixedsize = true,
-        width = 2, color=blue] // sets as circles
-    tab7 [label = '@@7']
-
-    node [shape = circle,
-        fixedsize = true,
-        width = 2, color=red] // sets as circles
-    tab5 [label = '@@5']
-
-    # several 'edge' statements
- tab1->tab2->tab3->tab4->tab6->tab7 [color=blue]
- tab1->tab8 [color=red,style=dashed]
-  tab2->tab9 [color=red,style=dashed]
-   tab3->tab10 [color=red,style=dashed]
-    tab4->tab11 [color=red,style=dashed]
-    tab8->tab5[color=red,style=dashed]
-    tab9->tab5[color=red,style=dashed]
-    tab10->tab5[color=red,style=dashed]
-    tab11->tab5[color=red,style=dashed]
-    tab6->tab1 [style=dashed]
-        tab6->tab2 [style=dashed]
-    tab6->tab5 [color=red, style=dashed]
-
- }
-      [1]: 'Questions relevant to stakeholders'
-      [2]: 'Appropriate design & methods'
-      [3]: 'Unbiased reporting'
-      [4]: 'Accessible full publication'
-      [5]: 'Wasteful research'
-      [6]: 'Evidence synthesis'
-      [7]: 'Utilisable research' 
-      [8]: 'Previous knowledge not accounted for'
-      [9]: 'Poor design; \\nquestionable research practices'
-      [10]: 'Lack of transparency'
-      [11]: 'Publications not available to practitioners \\n& decision makers'
-  ", height=1000)
-
-```
-
-```{r Randomise order of the estimates in the same study,echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Cumulative forest plot of the meta-analysis of Darras et al.
-(2018) on the difference between human observers and acoustic recorders in terms of species richness.
-Estimates from the same study inputted in random order\\label{forestX}"}
-set.seed(8955)
-metax<-meta1
-metax$order<-order(metax$year, runif(length(metax$year))) 
-res2 <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=metax, slab=paste(metax$first_author, metax$year, sep=","))
-cres2<-cumul(res2, order=order(metax$order))
-par(mfrow=c(1,2))
-forest(cres2)
-forest(cres)
+#ggsave("C:/Users/matthew.grainger/OneDrive - NINA/Papers_in_prep/Cumulative_meta_analysis/Research_waste/Figures/Fig1._Cumulative_Forest.png", p, device = "png")
 
@@
 Table 1 can be found at: https://github.com/DrMattG/Research_waste/blob/master/Manuscript/Table1.pdf
-
-```{r, echo=FALSE}
-
-# df<-data.frame("Research Stage"=c("Questions relevant to stakeholders","Questions relevant to stakeholders","Appropriate design and methods","Appropriate design and methods",
-#                                   "Appropriate design and methods","Appropriate design and methods","Unbiased reporting","Unbiased reporting","Unbiased reporting","Unbiased reporting",
-#                                   "Unbiased reporting","Accessible full publication","Evidence synthesis"),
-#                "Examples of potential for research waste"=c("Irrelevant questions asked","Previous knowledge not properly taken into account","Study poorly designed, under-powered (or over-powered.
-etc.)",
-#                                                             "Using inappropriate statistical tools (including overfitting etc.)","Questionable research practices3 lead to poor quality research","Questionable research practices3 lead to poor quality research",
-#                                                             "Lack of open data","Hypothesising after the results are known","p-hacking","File-drawer syndrome (only some studies are published)","Incomplete reporting, making evidence synthesis difficult or impossible",
-#                                                             "Publications not available to practitioners and decision makers","Research not designed or presented in the context of the existing knowledge"),
-#                "Where ecology and conservation can reduce waste"=c("Co-development of research questions with stakeholders and using appropriate methodology such as Delphi exercises to avoid issues such as group think or not including the right group of experts or stakeholders",
-#                "Make use of evidence synthesis methods (e.g.
-cumulative meta-analysis, systematic mapping, systematic reviews, meta-analysis) to identify questions that are not satisfactorily answered",
-#                "Use simulations or power-analysis prior to undertaking data collection.
-Predefine effect size of interest with stakeholders (i.e.
-do not rely on rules of thumb for “statistical significance”)",
-#                "Better training of early-career researchers in methods. Open code and data to ensure reproducibility of methods",
-#                "Open science (open methods and data, reproducible methods, sharing code, etc.)",
-#                "Better training of early-career researchers in methods of open science and evidence synthesis",
-#                "Open science (open methods and data, reproducible methods, sharing code, etc.)",
-#                "Pre-registration of hypotheses",
-#                "Open science (open methods and data, reproducible methods, sharing code, etc.)",
-#                "Pre-registration of hypotheses and methods. Open publishing (including preprints)",
-#                "Increasing knowledge of researchers and peer reviewers on what is essential to report, and changing journal guidelines where necessary to ensure all relevant information is reported",
-#                "Open access publishing, including making resources available to researchers to be able to publish open access","Using systematic reviews, systematic maps, meta-analysis, etc.
-to shape research priorities.
-Where good quality evidence is available these should be synthesised providing evidence to relevant stakeholders.
-Research gaps should be the focus of primary studies."
-#                ))
-# # # kable(df, "latex", booktabs = T, align = "c") %>% column_spec(1, bold=T) %>% collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle")
-# library(kableExtra)
-# #knitr::kable(df)
-# df %>% kable("latex") %>% as_image()
-
-```

### 2020-01-27 Manuscript/SuppMat.Rmd
2nd round review update 2

@@
 
-## Supplementary Information
+## Supplementary Methods
 
-# Methods
 We extracted the data and R code from the supplementary information in [@darras2018comparing] to recreate their analysis.
@@
 
+```{r flowc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "The production of research flows through five stages (blue lines) all of which can lead to research waste (red dashed lines).
+Ecology and conservation have begun to reduce waste by focusing on methodological improvements and open science.
+Evidence synthesis (including reporting to decision makers) can contribute to the reduction in research waste by influencing question setting and appropriate methods and design (black dashed lines).
+Poor evidence synthesis can also lead to research waste.
+\\label{flow}"}
+
+library(DiagrammeR)
+
+graph<-grViz("
+    digraph boxes_and_circles {
+    # a 'graph' statement
+    graph [overlap = true, fontsize = 10]
+
+    # left 'node' statements
+    node [shape = box, fontname = Helvetica]
+      tab1 [label = '@@1',color = blue]
+      tab2 [label = '@@2',color = blue]
+      tab3 [label = '@@3',color = blue]
+      tab4 [label = '@@4',color = blue]
+      tab6 [label = '@@6',color = blue]
+
+    # right node statements
+      tab8 [label = '@@8',color = red]
+      tab9 [label = '@@9',color = red]
+      tab10 [label = 'Lack of transparency',color = red]
+      tab11 [label = 'Publications not available to practitioners \\n& decision makers',color = red]
+
+      node [shape = circle,fixedsize = true,
+        width = 2, color=blue] // sets as circles
+    tab7 [label = '@@7']
+
+    node [shape = circle,
+        fixedsize = true,
+        width = 2, color=red] // sets as circles
+    tab5 [label = '@@5']
+
+        # several 'edge' statements
+ tab1->tab2->tab3->tab4->tab6->tab7 [color=blue]
+ tab1->tab8 [color=red,style=dashed]
+  tab2->tab9 [color=red,style=dashed]
+   tab3->tab10 [color=red,style=dashed]
+    tab4->tab11 [color=red,style=dashed]
+    tab8->tab5 [color=red,style=dashed]
+    tab9->tab5 [color=red,style=dashed]
+    tab10->tab5 [color=red,style=dashed]
+    tab11->tab5 [color=red,style=dashed]
+    tab6->tab1 [style=dashed]
+        tab6->tab2 [style=dashed]
+    tab6->tab5 [color=red, style=dashed]
+
+    tab8->tab9[style=invis]
+    tab9->tab10[style=invis]
+    tab10->tab11[style=invis]
+
+        # define ranks
+
+         subgraph {
+            rank = same; tab6; tab11
+         }
+
+    }
+       [1]: 'Questions are relevant to stakeholders'
+      [2]: 'Appropriate design & methods used'
+      [3]: 'Unbiased reporting'
+      [4]: 'Accessible full publication'
+      [5]: 'Wasteful research'
+      [6]: 'Evidence synthesis'
+      [7]: 'Utilisable research' 
+      [8]: 'Previous knowledge not accounted for'
+      [9]: 'Poor design; \\nquestionable research practices'
+        ")#, height=1000)
+
+graph
+
+```
+
 # References {-}

### 2020-01-27 Flowchart.R

@@ -82,8 +82,8 @@ digraph boxes_and_circles {
       tab6 [label = '@@6',color = blue]
       tab8 [label = '@@8',color = red]
       tab9 [label = '@@9',color = red]
-      tab10[label = '@@10',color = red]
-      tab11[label = '@@11',color = red]
+      tab10 [label = '@@10',color = red]
+      tab11 [label = '@@11',color = red]
       
   node [shape = circle,
         fixedsize = true,
@@ -123,3 +123,72 @@ digraph boxes_and_circles {
       [10]: 'Lack of transparency'
       [11]: 'Publications not available to practitioners \\n& decision makers'
   ", height=1000)
+
+
+
+grViz("
+    digraph boxes_and_circles {
+    # a 'graph' statement
+    graph [overlap = true, fontsize = 10]
+    
+    # left 'node' statements
+    node [shape = box, fontname = Helvetica]
+      tab1 [label = '@@1',color = blue]
+      tab2 [label = '@@2',color = blue]
+      tab3 [label = '@@3',color = blue]
+      tab4 [label = '@@4',color = blue]
+      tab6 [label = '@@6',color = blue]
+    
+    # right node statements
+      tab8 [label = '@@8',color = red]
+      tab9 [label = '@@9',color = red]
+      tab10 [label = 'Lack of transparency',color = red]
+      tab11 [label = 'Publications not available to practitioners \\n& decision makers',color = red]
+      
+      node [shape = circle,fixedsize = true,
+        width = 2, color=blue] // sets as circles
+    tab7 [label = '@@7']
+    
+    node [shape = circle,
+        fixedsize = true,
+        width = 2, color=red] // sets as circles
+    tab5 [label = '@@5']
+
+        # several 'edge' statements
+ tab1->tab2->tab3->tab4->tab6->tab7 [color=blue]
+ tab1->tab8 [color=red,style=dashed]
+  tab2->tab9 [color=red,style=dashed]
+   tab3->tab10 [color=red,style=dashed]
+    tab4->tab11 [color=red,style=dashed]
+    tab8->tab5 [color=red,style=dashed]
+    tab9->tab5 [color=red,style=dashed]
+    tab10->tab5 [color=red,style=dashed]
+    tab11->tab5 [color=red,style=dashed]
+    tab6->tab1 [style=dashed]
+        tab6->tab2 [style=dashed]
+    tab6->tab5 [color=red, style=dashed]
+    
+    tab8->tab9[style=invis]
[…]

### 2020-01-27 Manuscript/SuppMat.Rmd
2nd round review update

@@
 
-graph<-grViz("
+grViz("
     digraph boxes_and_circles {
@@
       [9]: 'Poor design; \\nquestionable research practices'
-        ")#, height=1000)
-
-graph
+        ", height=1000)
 

### 2020-01-27 Manuscript/Table1.Rmd
2nd round review update

--- 2019-12-02 Table1.Rmd
@@
 etc.)",
-                                                            "Using inappropriate statistical tools (including overfitting etc.)","Questionable research practices3 lead to poor quality research","Questionable research practices3 lead to poor quality research",
+                                                            "Using inappropriate statistical tools (including overfitting etc.)","Questionable research practices lead to poor quality research","Questionable research practices lead to poor quality research",
                                                             "Lack of open data","Hypothesising after the results are known","p-hacking","File-drawer syndrome (only some studies are published)","Incomplete reporting, making evidence synthesis difficult or impossible",
@@
 The research process stages (adapted from Chalmers & Glasziou, 2009), examples of potential research waste and how ecology and conservation can limit these.",border_left = T, border_right = T) %>% column_spec(1, bold=T) %>% collapse_rows(columns = 1:3, latex_hline = "major", valign = "middle") %>% kable_styling()
-#save_kable(k, "C:/Users/matthew.grainger/OneDrive - NINA/Papers_in_prep/Cumulative_meta_analysis/Research_waste/Mansuscript/Table1.pdf")
+#save_kable(k, "table1.pdf")
 

### 2020-01-27 Manuscript/MS.Rmd
2nd round review update
Resolve conflict

@@
 output:
-  pdf_document:
+  html_document: 
     fig_caption: yes
@@
     fig_caption: yes
-  html_document: 
+  pdf_document:
     fig_caption: yes
@@
 ## Outlook
+
+Research waste can be reduced and it is the responsibility of research funders, research institutions as well as individual researchers to do so.
+Researchers and funders could search for existing research syntheses in the literature and on synthesis platforms (e.g.
+https://www.conservationevidence.com/).
+We agree with the statement targeted at medicine 25 years ago that “We need...better research, and research done for the right reasons” [@altman1994scandal].
+Without a change in focus ecology and conservation funding will continue to be wasted which will be detrimental to our efforts to provide solutions to global societal challenges.
+
+### Box 1 Using cumulative meta-analysis to make research decisions
+
+=======
 

### 2020-01-27 Manuscript/SuppMat.Rmd
update 27.01.2020

--- 2020-01-27 Manuscript/SuppMat.Rmd
@@
     }
-       [1]: 'Questions are relevant to stakeholders'
-      [2]: 'Appropriate design & methods used'
-      [3]: 'Unbiased reporting'
+       [1]: 'Research question'
+      [2]: 'Study design & methods'
+      [3]: 'Reporting'
       [4]: 'Accessible full publication'

### 2020-01-27 Flowchart.R

@@ -180,9 +180,9 @@ grViz("
          
         
     }
-       [1]: 'Questions are relevant to stakeholders'
-      [2]: 'Appropriate design & methods used'
-      [3]: 'Unbiased reporting'
+       [1]: 'Research question'
+      [2]: 'Study design & methods'
+      [3]: 'Reporting'
       [4]: 'Accessible full publication'
       [5]: 'Wasteful research'
       [6]: 'Evidence synthesis'

### 2020-01-27 Rcode.R

@@ -93,69 +93,69 @@ RMA_alpha_identical=rma.mv(yi=alpha_ROM_identical,V=alpha_variance_ROM_identical
 res <- rma(RMA_alpha_identical$yi, RMA_alpha_identical$vi, data=meta1, slab=paste(meta1$first_author, meta1$year, sep=", "))
 cres<-cumul(res, order=order(meta1$year))
 
-meta1$IF<-c(2.27,0.967,0.967,0.967,1.29,0.41,0.967,0.967,0.967,0,0.538,2.055,2.53,2.53,0.622,1.92,1.895,0.94,1.895,0,0.967,1.75,6.36,0,0.41,0,0.41)
-cres2<-cumul(res, order=order(meta1$IF))
+#meta1$IF<-c(2.27,0.967,0.967,0.967,1.29,0.41,0.967,0.967,0.967,0,0.538,2.055,2.53,2.53,0.622,1.92,1.895,0.94,1.895,0,0.967,1.75,6.36,0,0.41,0,0.41)
+#cres2<-cumul(res, order=order(meta1$IF))
 
 #plot.cumul.rma(cres)
 
 
 #Figure1
-png("Figures/Figure1.png", width=850, height =400)
+#png("Figures/Figure1.png", width=850, height =400)
 forest(cres, col="blue")
 dev.off()
 
-forest(cres2, col="blue")
+#forest(cres2, col="blue")
 #Cumulative data frame
-cumzval<-cumsum(cres$zval)
-cum_sample<-cumsum(meta1$alpha_point_n_identical)  
-Year=meta1$year
+#cumzval<-cumsum(cres$zval)
+#cum_sample<-cumsum(meta1$alpha_point_n_identical)  
+#Year=meta1$year
 
-Year=gsub("a","",Year)
-Year=gsub("b","",Year)
-Year<-as.integer(Year)
-Year<-Year[order(Year)]
+#Year=gsub("a","",Year)
+#Year=gsub("b","",Year)
+#Year<-as.integer(Year)
+#Year<-Year[order(Year)]
 
-cum_df<-data.frame("z-score"=cumzval, "Cumulative n"=cum_sample, "Year"=Year)
+#cum_df<-data.frame("z-score"=cumzval, "Cumulative n"=cum_sample, "Year"=Year)
 #Labelling dataframe
-cum_df2<-cum_df[c(1,18,27),]
-png("Figures/Figure2.png", width=850, height =400)
-cum_df %>% 
[…]

### 2020-01-27 Flowchart.R
27.01.2020 Add Figs

@@ -191,4 +191,3 @@ grViz("
       [9]: 'Poor design; \\nquestionable research practices'
         ")
 #, height=1000)
-

### 2020-01-28 Manuscript/MS.Rmd
update 27.01.2020
ggsave path changed

--- 2020-01-27 Manuscript/MS.Rmd
@@
 
-#ggsave("C:/Users/matthew.grainger/OneDrive - NINA/Papers_in_prep/Cumulative_meta_analysis/Research_waste/Figures/Fig1._Cumulative_Forest.png", p, device = "png")
+#ggsave("C:/Users/matthew.grainger/OneDrive - NINA/Papers_in_prep/Cumulative_meta_analysis/Research_waste/Figures/Fig1._Cumulative_Forest.pdf", p, device = "pdf")
 
